---
title: "Age-at-risk stratified Cox PH analysis plots"
output: html_document
date: "`r Sys.Date()`"
---

# Age-at-risk stratified HRs between baseline frailty and outcomes

Cox proportional hazards models were adjusted for age-at-risk (5 year groups), sex (when appropriate), region, income level, marital status, occupation, education level, hypertension diagnosis, diabetes diagnosis, alcohol category, smoking category, fat mass (kg), poor diet, and physical activity.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Libraries

```{r}
library(ckbplotr)
library(cowplot)
library(stringr)
library(tidyverse)
```

## Open HRs 

HRs between baseline frailty and outcomes were stored in an RDS object (nested list structure)

```{r}
if (fi_type == "Original FI"){
  
  hrlist = readRDS(file = "J:/R projects/PhD-by-chapter/Chapter6-prospective-association/rds objects/output data/HRs baseline FI (original) and outcomes.rds")
  
  ylimit = c(0.85, 4)
  ybreak = c(1, 1.5, 2.5, 4)
  case_y_pos = 3.8
  hr_y_pos = 3.2
  
} else if (fi_type == "mFI-1"){
 
  hrlist = readRDS(file = "J:/R projects/PhD-by-chapter/Chapter6-prospective-association/rds objects/output data/HRs baseline FI (mFI-1) and outcomes.rds")
  
  ylimit = c(0.85, 2.5)
  ybreak = c(1, 1.5, 2.5)
  case_y_pos = 2.4
  hr_y_pos = 2.1
  
} else if (fi_type == "mFI-3"){
 
  hrlist = readRDS(file = "J:/R projects/PhD-by-chapter/Chapter6-prospective-association/rds objects/output data/HRs baseline FI (mFI-3) and outcomes.rds")
  
  ylimit = c(0.85, 2)
  ybreak = c(1, 1.5, 2)
  case_y_pos = 1.95
  hr_y_pos = 1.75
  
} else {
   
}
```

Nested within each of the 6 list elements:

```{r}
names(hrlist)
```

Are HRs between baseline frailty and the following 23 outcomes:

```{r}
names(hrlist[[1]])
```
Rename outcomes 

```{r}
# Save ordering of outcome
outcome_name = names(hrlist[[1]])

outcome_name = outcome_name %>% 
  str_replace(., "fatal", "(fatal)") %>% 
  str_replace(., "incidence", "(non-fatal)") %>% 
  str_replace(., "Fatal CVD (total)", "CVD mortality") %>%   
  str_replace(., "total", "(any)") %>% 
  str_replace(., "All cause mortality", "All-cause mortality")

outcome_name

# Rename list elements
for (i in 1:length(hrlist)){
  names(hrlist[[i]]) = outcome_name
}
```

## Create table of HRs with number of events

By levels of frailty (non-frail, pre-frail, and frail) and per 1SD increase

```{r}
# Frailty levels
func_get_hrs = function(df){ # create function to exponentiate HR estimates 
  
  df = df %>% 
   mutate(hr = format(round(exp(est), 2), nsmall=2),
          lq = format(round(exp(est-1.96*se), 2), nsmall=2),
          uq = format(round(exp(est+1.96*se), 2), nsmall=2)) %>% 
   mutate(hrs = paste0(hr, " (", lq, "-", uq, ")")) %>% 
   select(rf, n, hrs)
  
  return(df)
  
} 

hrs_filevel_total = lapply(hrlist[["HRs FI level and outcomes overall"]], func_get_hrs) # apply function to each element of list

hrs_filevel_total = data.table::rbindlist(hrs_filevel_total, idcol = "outcome") # combine data from each element of list

# Frailty per 1SD

hrs_fi1sd_total = data.table::rbindlist(hrlist[["HRs FI 1SD and outcomes overall"]], idcol = "outcome") %>% 
   mutate(hr = format(round(estimate, 2), nsmall=2),
          lq = format(round(conf.low, 2), nsmall=2),
          uq = format(round(conf.high, 2), nsmall=2)) %>% 
  mutate(hrs = paste0(hr, " (", lq, "-", uq, ")")) %>% 
  select(outcome, rf=term, n, hrs)

# Save ordering of outcome
outcome_order = pull(hrs_fi1sd_total, outcome)

# Combine data
hrs_total = bind_rows(hrs_filevel_total, hrs_fi1sd_total) %>% 
  mutate(outcome = factor(outcome, levels = outcome_order)) %>% 
  arrange(outcome)

hrs_total

write.csv(hrs_total, file = paste0("J:/R projects/PhD-by-chapter/Chapter6-prospective-association/outputs/age-at-risk stratified HRs/", fi_type, "/HRs_", fi_type, "_outcomes.csv"))
```

## Create figures 

### Baseline FI quintiles and outcomes

```{r, fig.width=10, fig.height=10}
plots1 = list()

# Create plots for each outcome

for (i in 1:length(hrlist[[1]])){ # 1 refers to list element with "HRs FI quintile and outcomes overall"
  
  # Get title for plot
  
  title = names(hrlist[[1]][i]) 
  
  # Create plot
  
  plots1[[i]] = ckbplotr::make_shape_plot(hrlist[[1]][[i]],
                            col.x          = "rf",
                            col.estimate   = "est",
                            col.stderr     = "se",
                            col.n          = "n",
                            exponentiate   = TRUE, 
                            xlims          = c(0.8, 5),
                            ylims          = ylimit,
                            ybreaks        = ybreak,
                            xlab           = paste0(fi_type, " quintile"),
                            ylab           = "Hazard ratio (95% CI)",
                            scalepoints    = TRUE,
                            title          = title,
                            printplot      = FALSE, 
                            showcode = FALSE)$plot

}

# Add number of events and HRs per 1SD increase in FI for each outcome

add_text_to_plot = function(i){
  
  # Get number of events and HRs for 1SD increase in FI
  
  tot = scales::label_comma()(
    sum(hrlist[[1]][[i]]$n, na.rm =TRUE)
  )

  hrs = hrlist[["HRs FI 1SD and outcomes overall"]][[i]] %>% # 7 refers to list element with "HRs FI 1SD and outcomes overall"
    mutate(hr = format(round(estimate, 2), nsmall=2),
          lq = format(round(conf.low, 2), nsmall=2),
          uq = format(round(conf.high, 2), nsmall=2)) %>% 
    mutate(hrs = paste0(hr, " (", lq, "-", uq, ")")) %>% 
    pull(hrs)
  
  # Create plot
  
  output = plots1[[i]] +
      # Plot n events text and HRs per 1SD increase in FI
      geom_text(aes(y = case_y_pos,
                    x = 3,
                    label = paste0("(", tot, " cases)")),
                size  = 3.5) +
      geom_text(aes(y = hr_y_pos,
                    x = 3,
                    label = paste0("HR (95%CI): ", hrs, "\nper 1SD increase")),
                size  = 3.5)
  
  return(output)

}

plots = list()

plots = lapply(c(1:23), add_text_to_plot)

plot_stroke = cowplot::plot_grid(plots[[1]], plots[[2]], plots[[3]], 
                                 plots[[4]], plots[[5]], plots[[6]],
                                 plots[[7]], plots[[8]], plots[[9]],
                                 ncol=3, nrow=3, align = "hv", scale = 0.9)

plot_ihd = cowplot::plot_grid(plots[[10]], plots[[11]], plots[[12]], 
                              plots[[13]], plots[[14]], plots[[15]],
                              plots[[19]], plots[[20]], plots[[21]],
                              ncol=3, nrow=3, align = "hv", scale = 0.9)

plot_total = cowplot::plot_grid(plots[[22]], plots[[23]],
                            ncol=2, nrow=1, align = "hv", scale = 0.9)

plot_stroke
plot_ihd
plot_total

cowplot::save_plot(paste0("J:/R projects/PhD-by-chapter/Chapter6-prospective-association/outputs/age-at-risk stratified HRs/", fi_type, "/HRplots_", fi_type, "_quintile_stroke.png"), plot_stroke, base_height = 13, base_asp = 0.8) # base_asp = width/height
cowplot::save_plot(paste0("J:/R projects/PhD-by-chapter/Chapter6-prospective-association/outputs/age-at-risk stratified HRs/", fi_type, "/HRplots_", fi_type, "_quintile_ihd.png"), plot_ihd, base_height = 13, base_asp = 0.8) # base_asp = width/height
cowplot::save_plot(paste0("J:/R projects/PhD-by-chapter/Chapter6-prospective-association/outputs/age-at-risk stratified HRs/", fi_type, "/HRplots_", fi_type, "_quintile_mortality.png"), plot_total, base_height = 6, base_asp = 1.5) # base_asp = width/height
```


```{r}
knitr::knit_exit()
```

### Baseline FI levels and outcomes

```{r, fig.width=10, fig.height=10}
plots = list()

for (i in 1:length(hrlist[[4]])){
  
  title = names(hrlist[[4]][i]) 
  
  hrlist[[4]][[i]]$rf = factor(hrlist[[4]][[i]]$rf, levels = c("Non-frail", "Pre-frail", "Frail"))
  
  plots[[i]] = ckbplotr::make_shape_plot(hrlist[[4]][[i]],
                            col.x          = "rf",
                            col.estimate   = "est",
                            col.stderr     = "se",
                            col.n          = "n",
                            exponentiate   = TRUE, 
                            xlims          = c(1, 3),
                            ylims          = ylimit,
                            ybreaks        = ybreak,
                            xlab           = "",
                            ylab           = "Hazard ratio (95% CI)",
                            scalepoints    = TRUE,
                            title          = title,
                            printplot      = FALSE, 
                            showcode = FALSE)
  
}

plot2a = cowplot::plot_grid(plots[[1]]$plot, plots[[2]]$plot, plots[[3]]$plot, 
                            plots[[4]]$plot, plots[[5]]$plot, plots[[6]]$plot,
                            plots[[7]]$plot, plots[[8]]$plot, plots[[9]]$plot,
                            ncol=3, nrow=3, align = "hv")

plot2b = cowplot::plot_grid(plots[[10]]$plot, plots[[11]]$plot, plots[[12]]$plot, 
                            plots[[13]]$plot, plots[[14]]$plot, plots[[15]]$plot,
                            plots[[22]]$plot, plots[[23]]$plot,
                            ncol=3, nrow=3, align = "hv")

plot2c = cowplot::plot_grid(plots[[16]]$plot, plots[[17]]$plot, plots[[18]]$plot, 
                            plots[[19]]$plot, plots[[20]]$plot, plots[[21]]$plot,
                            ncol=3, nrow=3, align = "hv")

plot2a
plot2b
plot2c

cowplot::save_plot(paste0("J:/R projects/PhD-by-chapter/Chapter6-prospective-association/outputs/age-at-risk stratified HRs/", fi_type, " FI/HRplots_FI", fi_type, "_level_CVD1_fulladj.png"), plot2a, base_height = 8, base_aspect_ratio = 1)
cowplot::save_plot(paste0("J:/R projects/PhD-by-chapter/Chapter6-prospective-association/outputs/age-at-risk stratified HRs/", fi_type, " FI/HRplots_FI", fi_type, "_level_CVD2_fulladj.png"), plot2b, base_height = 8, base_aspect_ratio = 1)
cowplot::save_plot(paste0("J:/R projects/PhD-by-chapter/Chapter6-prospective-association/outputs/age-at-risk stratified HRs/", fi_type, " FI/HRplots_FI", fi_type, "_level_CVD3_fulladj.png"), plot2c, base_height = 8, base_aspect_ratio = 1)
```

### HRs by sex 

Combine data for male and female into one data frame

```{r}
# Add 'sex' variable to all dataframes in relevant list
hrlist[[2]] = map2(hrlist[[2]], "Male", ~cbind(.x, Sex = .y)) # map2 iterates a function over multiple arguments simultaenously. Each input is processed in parallel with others
hrlist[[3]] = map2(hrlist[[3]], "Female", ~cbind(.x, Sex = .y))
hrlist[[5]] = map2(hrlist[[5]], "Male", ~cbind(.x, Sex = .y)) 
hrlist[[6]] = map2(hrlist[[6]], "Female", ~cbind(.x, Sex = .y))

# Append the two lists 
temp = append(hrlist[[2]], hrlist[[3]])
# Bind rows for dataframes in the list with the same outcome
hrlist_sex_quintile = tapply(temp, names(temp), dplyr::bind_rows) # tapply applies a function to each cell; grouping variable (names(temp)) is used to apply functions by this grouping

temp = append(hrlist[[5]], hrlist[[6]])
# Bind rows for dataframes in the list with the same outcome
hrlist_sex_level = tapply(temp, names(temp), dplyr::bind_rows) # tapply applies a function to each cell; grouping variable (names(temp)) is used to apply functions by this grouping

# Reorder list elements
hrlist_sex_quintile = hrlist_sex_quintile[c("Stroke incidence", "Stroke fatal", "Stroke total",
                                            "Ischaemic stroke incidence", "Ischaemic stroke fatal", "Ischaemic stroke total",
                                            "Haemorrhagic stroke incidence", "Haemorrhagic stroke fatal", "Haemorrhagic stroke total",
                                            "IHD incidence", "IHD fatal", "IHD total",
                                            "MI incidence", "MI fatal", "MI total",
                                            "MVE incidence", "MVE fatal", "MVE total",
                                            "MCE incidence", "MCE fatal", "MCE total",
                                            "Fatal CVD total", "All cause mortality")]


hrlist_sex_level = hrlist_sex_level[c("Stroke incidence", "Stroke fatal", "Stroke total",
                                      "Ischaemic stroke incidence", "Ischaemic stroke fatal", "Ischaemic stroke total",
                                      "Haemorrhagic stroke incidence", "Haemorrhagic stroke fatal", "Haemorrhagic stroke total",
                                      "IHD incidence", "IHD fatal", "IHD total",
                                      "MI incidence", "MI fatal", "MI total",
                                      "MVE incidence", "MVE fatal", "MVE total",
                                      "MCE incidence", "MCE fatal", "MCE total",
                                      "Fatal CVD total", "All cause mortality")]
```

```{r, eval=FALSE, include=FALSE}
# Testing code for ckbplotr

# ckbplotr::make_shape_plot(hrlist_sex_quintile[[1]],
#                             col.x          = "rf",
#                             col.estimate   = "est",
#                             col.stderr     = "se",
#                             col.n          = "n",
#                             col.group      = "Sex",
#                             exponentiate   = TRUE, 
#                             xlims          = c(1, 5),
#                             ylims          = c(0.85, 2.5),
#                             xlab           = "",
#                             ylab           = "Hazard ratio (95% CI)",
#                             scalepoints    = TRUE,
#                             title          = "",
#                             printplot      = TRUE, 
#                             showcode = TRUE)

# Create the ggplot
plot <- ggplot(data = hrlist_sex_quintile[[1]],
               aes(x = rf, y = exp(est), group = as.factor(Sex))) +

  # Plot the point estimates
  geom_point(aes(size = 1/se,
                 fill = as.factor(Sex)),
             shape = 22,
             colour = "black",
             stroke = 0.5,
             position = position_dodge(width=0.8)) +
  
  # Plot point estimates text
  geom_text(aes(y = exp(est+1.96*se),
                label = format(round(exp(est), 2), nsmall = 2)),
            vjust = -0.8,
            size  = 3,
             position = position_dodge(width=0.8)) +
  
  # Plot n events text
  geom_text(aes(y = exp(est-1.96*se),
                label = n),
            vjust = 1.8,
            size  = 3,
             position = position_dodge(width=0.8)) +
  
  # Plot the CIs
  geom_linerange(aes(ymin = exp(est-1.96*se),
                     ymax = exp(est+1.96*se)),
                 colour = "black",
                 lwd = 0.5,
             position = position_dodge(width=0.8)) +
  
  # Set the scale for the size of boxes
  scale_radius(guide  = "none",
               limits = c(0, NA_real_),
               range  = c(0, 3)) +
  
  # Use identity for aesthetic scales
  scale_shape_identity() +
  scale_colour_identity() +
  
  # Set the scale for fill colours
  scale_fill_grey(start = 0, end = 1, guide = "none") +
  
  # Set the y-axis scale
  scale_y_continuous(trans = "log") +
  
  # Add titles
  xlab("") +
  ylab("Hazard ratio (95% CI)")
  
# Plot like a CKB plot
test = ckbplotr::plot_like_ckb(plot    = plot,
                        xlims          = c(1, 5),
                        ylims          = ylimit,
                        gap            = c(0.1, 0.025),
                        ext            = c(0.025, 0.025),
                        ratio          = 1.5,
                        base_size      = 11,
                        base_line_size = 0.5)

test
```

#### Baseline FI quintiles and outcomes by sex

```{r, fig.width=12, fig.height=10}
plots = list()

for (i in 1:length(hrlist_sex_quintile)){
  
  title = names(hrlist_sex_quintile[i]) 
  
  # Create the ggplot
  plot <- ggplot(data = hrlist_sex_quintile[[i]],
                 aes(x = rf, y = exp(est), group = as.factor(Sex))) +
  
    # Plot the point estimates
    geom_point(aes(size = 1/se,
                   fill = as.factor(Sex)),
                   shape = 22,
                   colour = "black",
                   stroke = 0.5,
                   position = position_dodge(width=0.8)) +
    
    # Plot point estimates text
    geom_text(aes(y = exp(est+1.96*se),
                  label = format(round(exp(est), 2), nsmall = 2)),
                  vjust = -0.8,
                  size  = 3,
                  position = position_dodge(width=0.8)) +
    
    # Plot n events text
    geom_text(aes(y = exp(est-1.96*se),
                  label = n),
                  vjust = 1.8,
                  size  = 3,
                  position = position_dodge(width=0.8)) +
    
    # Plot the CIs
    geom_linerange(aes(ymin = exp(est-1.96*se),
                       ymax = exp(est+1.96*se)),
                       colour = "black",
                       lwd = 0.5,
                       position = position_dodge(width=0.8)) +
    
    # Set the scale for the size of boxes
    scale_radius(guide  = "none",
                 limits = c(0, NA_real_),
                 range  = c(0, 3)) +
    
    # Use identity for aesthetic scales
    scale_shape_identity() +
    scale_colour_identity() +
    
    # Set the scale for fill colours
    scale_fill_grey(start = 0, end = 1, guide = "none") +
    
    # Set the y-axis scale
    scale_y_continuous(trans = "log",
                       breaks = ybreak) +
    
    # Add titles
    xlab("") +
    ylab("Hazard ratio (95% CI)") +
    ggtitle(title)
    
  # Plot like a CKB plot
  plots[[i]] = ckbplotr::plot_like_ckb(plot           = plot,
                                       xlims          = c(1, 5),
                                       ylims          = ylimit,
                                       gap            = c(0.1, 0.025),
                                       ext            = c(0.025, 0.025),
                                       ratio          = 1.5,
                                       base_size      = 11,
                                       base_line_size = 0.5)
  
}

plot3a = cowplot::plot_grid(plots[[1]], plots[[2]], plots[[3]], 
                            plots[[4]], plots[[5]], plots[[6]],
                            plots[[7]], plots[[8]], plots[[9]],
                            ncol=3, nrow=3, align = "hv")

plot3b = cowplot::plot_grid(plots[[10]], plots[[11]], plots[[12]], 
                            plots[[13]], plots[[14]], plots[[15]],
                            plots[[22]], plots[[23]],
                            ncol=3, nrow=3, align = "hv")

plot3c = cowplot::plot_grid(plots[[16]], plots[[17]], plots[[18]], 
                            plots[[19]], plots[[20]], plots[[21]],
                            ncol=3, nrow=3, align = "hv")

plot3a
plot3b
plot3c

cowplot::save_plot(paste0("J:/R projects/PhD-by-chapter/Chapter6-prospective-association/outputs/age-at-risk stratified HRs/", fi_type, " FI/HRplots_FI", fi_type, "_quintile_CVD1_fulladj_sex.png"), plot3a, base_height = 12, base_aspect_ratio = 1)
cowplot::save_plot(paste0("J:/R projects/PhD-by-chapter/Chapter6-prospective-association/outputs/age-at-risk stratified HRs/", fi_type, " FI/HRplots_FI", fi_type, "_quintile_CVD2_fulladj_sex.png"), plot3b, base_height = 12, base_aspect_ratio = 1)
cowplot::save_plot(paste0("J:/R projects/PhD-by-chapter/Chapter6-prospective-association/outputs/age-at-risk stratified HRs/", fi_type, " FI/HRplots_FI", fi_type, "_quintile_CVD3_fulladj_sex.png"), plot3c, base_height = 12, base_aspect_ratio = 1)
```

#### Baseline FI levels and outcomes by sex

```{r, fig.width=12, fig.height=10}
plots = list()

for (i in 1:length(hrlist_sex_level)){
  
  title = names(hrlist_sex_level[i]) 
  
  hrlist_sex_level[[i]]$rf = factor(hrlist_sex_level[[i]]$rf, levels = c("Non-frail", "Pre-frail", "Frail"))
  
  # Create the ggplot
  plot <- ggplot(data = hrlist_sex_level[[i]],
                 aes(x = rf, y = exp(est), group = as.factor(Sex))) +
  
    # Plot the point estimates
    geom_point(aes(size = 1/se,
                   fill = as.factor(Sex)),
                   shape = 22,
                   colour = "black",
                   stroke = 0.5,
                   position = position_dodge(width=0.8)) +
    
    # Plot point estimates text
    geom_text(aes(y = exp(est+1.96*se),
                  label = format(round(exp(est), 2), nsmall = 2)),
                  vjust = -0.8,
                  size  = 3,
                  position = position_dodge(width=0.8)) +
    
    # Plot n events text
    geom_text(aes(y = exp(est-1.96*se),
                  label = n),
                  vjust = 1.8,
                  size  = 3,
                  position = position_dodge(width=0.8)) +
    
    # Plot the CIs
    geom_linerange(aes(ymin = exp(est-1.96*se),
                       ymax = exp(est+1.96*se)),
                       colour = "black",
                       lwd = 0.5,
                       position = position_dodge(width=0.8)) +
    
    # Set the scale for the size of boxes
    scale_radius(guide  = "none",
                 limits = c(0, NA_real_),
                 range  = c(0, 3)) +
    
    # Use identity for aesthetic scales
    scale_shape_identity() +
    scale_colour_identity() +
    
    # Set the scale for fill colours
    scale_fill_grey(start = 0, end = 1, guide = "none") +
    
    # Set the y-axis scale
    scale_y_continuous(trans = "log",
                       breaks = ybreak) +
    
    # Add titles
    xlab("") +
    ylab("Hazard ratio (95% CI)") +
    ggtitle(title)
    
    # Plot like a CKB plot
    plots[[i]] = ckbplotr::plot_like_ckb(plot           = plot,
                                         xlims          = c(1, 3),
                                         ylims          = ylimit,
                                         gap            = c(0.180, 0.025),
                                         ext            = c(0.025, 0.025),
                                         ratio          = 1.5,
                                         base_size      = 11,
                                         base_line_size = 0.5)

}

plot4a = cowplot::plot_grid(plots[[1]], plots[[2]], plots[[3]], 
                            plots[[4]], plots[[5]], plots[[6]],
                            plots[[7]], plots[[8]], plots[[9]],
                            ncol=3, nrow=3, align = "hv")

plot4b = cowplot::plot_grid(plots[[10]], plots[[11]], plots[[12]], 
                            plots[[13]], plots[[14]], plots[[15]],
                            plots[[22]], plots[[23]],
                            ncol=3, nrow=3, align = "hv")

plot4c = cowplot::plot_grid(plots[[16]], plots[[17]], plots[[18]], 
                            plots[[19]], plots[[20]], plots[[21]],
                            ncol=3, nrow=3, align = "hv")

plot4a
plot4b
plot4c

cowplot::save_plot(paste0("J:/R projects/PhD-by-chapter/Chapter6-prospective-association/outputs/age-at-risk stratified HRs/", fi_type, " FI/HRplots_FI", fi_type, "_level_CVD1_fulladj_sex.png"), plot4a, base_height = 12, base_aspect_ratio = 1)
cowplot::save_plot(paste0("J:/R projects/PhD-by-chapter/Chapter6-prospective-association/outputs/age-at-risk stratified HRs/", fi_type, " FI/HRplots_FI", fi_type, "_level_CVD2_fulladj_sex.png"), plot4b, base_height = 12, base_aspect_ratio = 1)
cowplot::save_plot(paste0("J:/R projects/PhD-by-chapter/Chapter6-prospective-association/outputs/age-at-risk stratified HRs/", fi_type, " FI/HRplots_FI", fi_type, "_level_CVD3_fulladj_sex.png"), plot4c, base_height = 12, base_aspect_ratio = 1)
```