---
title: "Age-at-risk stratified Cox PH analysis plots"
output: html_document
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Libraries

```{r}
library(ckbplotr)
library(cowplot)
library(tidyverse)
```

## Open HRs 

HRs between baseline frailty and outcomes were stored in an RDS object (nested list structure)

```{r}
if (outcome == "Stroke" & fi_type == "mFI-1"){
  
  hrlist = readRDS(file = "J:/R projects/PhD-by-chapter/Chapter6-prospective-association/rds objects/output data/subgroup analysis/HRs mFI-1 and event_Stroke_total_subgroup.rds")
  
} else if (outcome == "IHD" & fi_type == "mFI-1"){
  
  hrlist = readRDS(file = "J:/R projects/PhD-by-chapter/Chapter6-prospective-association/rds objects/output data/subgroup analysis/HRs mFI-1 and event_IHD_total_subgroup.rds")
  
} else if (outcome == "CVD mortality" & fi_type == "mFI-1"){
  
  hrlist = readRDS(file = "J:/R projects/PhD-by-chapter/Chapter6-prospective-association/rds objects/output data/subgroup analysis/HRs mFI-1 and event_Fatal_CVD_total_subgroup.rds")
  
} else if (outcome == "All-cause mortality" & fi_type == "mFI-1"){
  
  hrlist = readRDS(file = "J:/R projects/PhD-by-chapter/Chapter6-prospective-association/rds objects/output data/subgroup analysis/HRs mFI-1 and event_All_cause_mortality_subgroup.rds")
  
} else if (outcome == "Stroke" & fi_type == "mFI-3"){
  
  hrlist = readRDS(file = "J:/R projects/PhD-by-chapter/Chapter6-prospective-association/rds objects/output data/subgroup analysis/HRs mFI-3 and event_Stroke_total_subgroup.rds")
  
} else if (outcome == "IHD" & fi_type == "mFI-3"){
  
  hrlist = readRDS(file = "J:/R projects/PhD-by-chapter/Chapter6-prospective-association/rds objects/output data/subgroup analysis/HRs mFI-3 and event_IHD_total_subgroup.rds")
  
} else if (outcome == "CVD mortality" & fi_type == "mFI-3"){
  
  hrlist = readRDS(file = "J:/R projects/PhD-by-chapter/Chapter6-prospective-association/rds objects/output data/subgroup analysis/HRs mFI-3 and event_Fatal_CVD_total_subgroup.rds")
  
} else if (outcome == "All-cause mortality" & fi_type == "mFI-3"){
  
  hrlist = readRDS(file = "J:/R projects/PhD-by-chapter/Chapter6-prospective-association/rds objects/output data/subgroup analysis/HRs mFI-3 and event_All_cause_mortality_subgroup.rds")
  
} else {}
```

Nested within each list element:

```{r}
str(hrlist)
```

HRs for outcome per 1SD increase in FI

```{r}
hrlist[[1]]
```

Tests of hetereogeneity and trend

```{r}
hrlist[[2]]
```
Open overall results

```{r}
if (fi_type == "mFI-1"){
  
  hrlist2 = readRDS(file = "J:/R projects/PhD-by-chapter/Chapter6-prospective-association/rds objects/output data/HRs baseline FI (mFI-1) and outcomes.rds")
  
  if (outcome == "Stroke"){
  
    hrlist2 = hrlist2[["HRs FI 1SD and outcomes overall"]][["Stroke total"]]
    
  } else if (outcome == "IHD"){
    
    hrlist2 = hrlist2[["HRs FI 1SD and outcomes overall"]][["IHD total"]]
    
  } else if (outcome == "CVD mortality"){
    
    hrlist2 = hrlist2[["HRs FI 1SD and outcomes overall"]][["Fatal CVD total"]]
    
  } else if (outcome == "All-cause mortality"){
    
    hrlist2 = hrlist2[["HRs FI 1SD and outcomes overall"]][["All cause mortality"]]
  } else {}
  
} else if (fi_type == "mFI-3"){
  
  hrlist2 = readRDS(file = "J:/R projects/PhD-by-chapter/Chapter6-prospective-association/rds objects/output data/HRs baseline FI (mFI-3) and outcomes.rds")
  
  if (outcome == "Stroke"){
  
    hrlist2 = hrlist2[["HRs FI 1SD and outcomes overall"]][["Stroke total"]]
    
  } else if (outcome == "IHD"){
    
    hrlist2 = hrlist2[["HRs FI 1SD and outcomes overall"]][["IHD total"]]
    
  } else if (outcome == "CVD mortality"){
    
    hrlist2 = hrlist2[["HRs FI 1SD and outcomes overall"]][["Fatal CVD total"]]
    
  } else if (outcome == "All-cause mortality"){
    
    hrlist2 = hrlist2[["HRs FI 1SD and outcomes overall"]][["All cause mortality"]]
  } else {}
  
} else {}

hrlist2 = hrlist2 %>% 
  select(estimate = beta, stderr = std.error, n) %>% 
  mutate(heading = "Overall",
         variable = "overall",
         subheading = "Total events")

hrlist2
```
## Create plots

Create data frame of estimates and test results

```{r}
results = hrlist[[1]] %>% 
  group_by(rf) %>% 
  mutate(id = row_number()) %>% 
  ungroup() %>% 
  mutate(variable = paste0(rf, id)) %>% 
  select(variable, subheading, label, estimate, stderr, n) %>% 
  mutate(subheading = factor(subheading, levels = c("Age group",
                                                    "Age groups",
                                                    "Sex",
                                                    "Region",
                                                    "Study site",
                                                    "Highest education",
                                                    "Highest education level",
                                                    "Income level",
                                                    "Alcohol category", "Smoking category", "Smoking category (quantity)",
                                                    "Fat mass quintile",
                                                    "Hypertension", "Systolic blood pressure", "SBP quintile",
                                                    "Diabetes", "Random glucose", "Random glucose quintile",
                                                    "Physical activity quintile"))) %>% 
  arrange(subheading) %>% 
  mutate(across(subheading, as.character),
         heading = "Subgroup")

results = bind_rows(hrlist2, results)

head(results)
```

```{r}
tests = hrlist[[2]] %>% 
  select(-rf) %>%
  mutate(subheading = case_when(subheading == "agegroup" ~ "Age group",
                                subheading == "agegroup2" ~ "Age groups",
                                subheading == "sex" ~ "Sex",
                                subheading == "region" ~ "Region",
                                subheading == "region2" ~ "Study site",
                                subheading == "education" ~ "Highest education level",
                                subheading == "education2" ~ "Highest education",
                                subheading == "income" ~ "Income level",
                                subheading == "fatmass" ~ "Fat mass quintile",
                                subheading == "alcohol" ~ "Alcohol category",
                                subheading == "smoking" ~ "Smoking category",
                                subheading == "smoking2" ~ "Smoking category (quantity)",
                                subheading == "hypertension" ~ "Hypertension",
                                subheading == "diabetes" ~ "Diabetes",
                                subheading == "sbp" ~ "Systolic blood pressure",
                                subheading == "glucose" ~ "Random glucose",
                                subheading == "sbp2" ~ "SBP quintile",
                                subheading == "glucose2" ~ "Random glucose quintile",
                                subheading == "met" ~ "Physical activity quintile")) %>% 
  left_join(select(results, heading, subheading, label, variable), by = c("subheading", "label"))

head(tests)
```

Subset results

```{r}
covlist1 = c("Sex", "Age group", "Region", "Highest education", "Fat mass quintile", "Alcohol category", "Smoking category", "Systolic blood pressure", "Random glucose", "Total events")

covlist2 = c("Age groups", "Income level", "Highest education level", "Smoking category (quantity)", "Hypertension", "SBP quintile", "Diabetes", "Random glucose quintile", "Total events")

covlist3 = c("Study site", "Total events")
```

```{r}
results1 = results %>% filter(subheading %in% covlist1)
results2 = results %>% filter(subheading %in% covlist2)
results3 = results %>% filter(subheading %in% covlist3)

tests1 = tests %>% filter(subheading %in% covlist1)
tests2 = tests %>% filter(subheading %in% covlist2)
tests3 = tests %>% filter(subheading %in% covlist3)
```

Create row labels

```{r}
row_labels = results[, c("heading", "subheading", "label", "variable")]
head(row_labels)

row_labels$label = stringr::str_replace(row_labels$label, ">=", as.character(expression("\u2265"))) # add greater than or equal to sign

row_labels1 = row_labels %>% filter(subheading %in% covlist1)
row_labels2 = row_labels %>% filter(subheading %in% covlist2)
row_labels3 = row_labels %>% filter(subheading %in% covlist3)
```

Set x-axis limits and breaks

```{r}
if (outcome == "Stroke"){
  
  xlim          = c(0.965, 1.21)
  xticks        = c(1, 1.1, 1.2)
  nullvall      = 1
                           
} else if (outcome == "IHD"){
  
  xlim          = c(1, 1.265)
  xticks        = c(1, 1.1, 1.2)
  nullvall      = 1
                           
} else if (outcome == "CVD mortality"){
  
  xlim          = c(1.1, 1.45)
  xticks        = c(1.1, 1.2, 1.3, 1.4)
  nullvall      = NULL
                           
} else if (outcome == "All-cause mortality"){

  xlim          = c(1.1, 1.4)
  xticks        = c(1.1, 1.3, 1.2, 1.4)
  nullvall      = NULL
  
} else {}
```

Plot

```{r, fig.height=10, fig.width=8}
plot1 = ckbplotr::make_forest_plot(panels        = list(results1),
                           col.key       = "variable",
                           exponentiate = TRUE,
                           row.labels    = row_labels1,
                           row.labels.levels = c("heading", "subheading", "label"),
                           rows          = c("Subgroup", "Overall"),
                           panel.names = c(outcome), # outcome
                           ci.delim      = "-",
                           xlim          = xlim,
                           xticks        = xticks,
                           nullval       = nullvall,  
                           blankrows     = c(1, 1, 0, 1),
                           scalepoints   = TRUE,
                           pointsize     = 3,
                           col.left      = c("n"),
                           col.left.heading = c("No. of\nevents"),
                           col.heading.space = 1.5,
                           bold.labels   = c("heading", "subheading"), 
                           addtext       = list(tests1),
                           right.space   = unit(35, "mm"))$plot +
  theme(plot.margin = unit(c(1,5,1,1), "cm")) # trbl 

plot1
```

```{r}
plot2 = ckbplotr::make_forest_plot(panels        = list(results2),
                           col.key       = "variable",
                           row.labels    = row_labels2,
                           row.labels.levels = c("heading", "subheading", "label"),
                           rows          = c("Subgroup", "Overall"),
                           exponentiate  = TRUE,
                           panel.names = c(outcome),
                           ci.delim      = "-",
                           xlim          = xlim,
                           xticks        = xticks,
                           nullval       = nullvall,
                           blankrows     = c(1, 1, 0, 1),
                           scalepoints   = TRUE,
                           pointsize     = 3,
                           col.left      = c("n"),
                           col.left.heading = c("No. of\nevents"),
                           col.heading.space = 1.5,
                           bold.labels   = c("heading", "subheading"), 
                           addtext       = list(tests2),
                           right.space   = unit(35, "mm"))$plot +
  theme(plot.margin = unit(c(1,5,1,1), "cm")) # trbl  

plot2
```

```{r}
plot3 = ckbplotr::make_forest_plot(panels        = list(results3),
                           col.key       = "variable",
                           row.labels    = row_labels3,
                           row.labels.levels = c("heading", "subheading", "label"),
                           rows          = c("Subgroup", "Overall"),
                           exponentiate  = TRUE,
                           panel.names = c(outcome),
                           ci.delim      = "-",
                           xlim          = xlim,
                           xticks        = xticks,
                           nullval       = nullvall,
                           blankrows     = c(1, 1, 0, 1),
                           scalepoints   = TRUE,
                           pointsize     = 3,
                           col.left      = c("n"),
                           col.left.heading = c("No. of\nevents"),
                           col.heading.space = 1.5,
                           bold.labels   = c("heading", "subheading"), 
                           addtext       = list(tests3),
                           right.space   = unit(35, "mm"))$plot +
  theme(plot.margin = unit(c(1,5,1,1), "cm")) # trbl  

plot3
```

Save plot

```{r}
cowplot::save_plot(paste0("J:/R projects/PhD-by-chapter/Chapter6-prospective-association/outputs/age-at-risk stratified HRs/subgroup/forest_", fi_type, "_", outcome, "1.png"), plot1, base_height = 10, base_asp = 0.7) # base_asp = width/height

cowplot::save_plot(paste0("J:/R projects/PhD-by-chapter/Chapter6-prospective-association/outputs/age-at-risk stratified HRs/subgroup/forest_", fi_type, "_", outcome, "2.png"), plot2, base_height = 10, base_asp = 0.7) # base_asp = width/height

cowplot::save_plot(paste0("J:/R projects/PhD-by-chapter/Chapter6-prospective-association/outputs/age-at-risk stratified HRs/subgroup/forest_", fi_type, "_", outcome, "3.png"), plot3, base_height = 5, base_asp = 1.5) # base_asp = width/height
```


```{r}
knitr::knit_exit()
```


Step 2: Remove the axis text for all but the first plot.

```{r}
p1 <- forestplot1$plot

p2 <- forestplot2$plot + 
  theme(axis.text.y = element_blank())
```

Step 3: Arrange the plots using gridExtra (there may be other packages that also work). Adjust widths until the plots are the desired widths in your final output file.

```{r, fig.height=10, fig.width=8}
final_plot = gridExtra::grid.arrange(p1, p2, nrow = 1, widths = c(1, 0.6))

cowplot::save_plot("J:/R projects/PhD-by-chapter/Chapter6-prospective-association/outputs/age-at-risk stratified HRs/modified FI/HRplots_subgroups_FImodified_1SD_IHDStroke_fulladj.png", final_plot, base_height = 10, base_aspect_ratio = 0.8) # aspect ratio = width/height
```

### Subgroup analysis for fat mass quintile

```{r}
# Heterogeneity and trend tests
test_het = as.data.frame(test_het_fatmass)
test_het$variable = "fatmass"
names(test_het) = c("het_stat", "het_p", "variable")

test_trend = as.data.frame(test_trend_fatmass)
test_trend$variable = "fatmass"
names(test_trend) = c("trend_stat", "trend_p", "variable")

results_extra = full_join(test_het, test_trend, by = "variable") %>% 
  mutate(het_p = format.pval(het_p, eps=.001, digits=1),
         trend_p = format.pval(trend_p, eps=.001, digits=1))

# HRs
final_results = bind_rows(data_fatmass)

final_results
```

```{r}
# Create group labels
final_results = final_results %>% 
  separate(variable, c("subheading", "label"), sep = "_", remove = FALSE) 

# Select the last group level of each subgroup to merge onto results_extra
labels_to_merge = final_results %>%
  group_by(subheading) %>% 
  dplyr::mutate(id = row_number()) %>% 
  filter(id == max(id)) %>% 
  select(variable, subheading)

results_extra = results_extra %>% 
  select(subheading = variable, everything()) %>% 
  full_join(labels_to_merge, by = "subheading") %>% # add labels
  select(-subheading) %>% 
  mutate_if(is.numeric, ~as.character(round(., 1))) %>% # round statistic to 1 dp
  mutate_at(vars(het_p, trend_p), ~if_else(grepl("^0", .), paste0("=", .), .)) # add = to pvalue

# Relabel headings 
final_results = final_results %>% 
  mutate(subheading = "Fat mass")
```

Save data for plotting later

```{r}
stroke_results = list(final_results, results_extra)
ihd_results = list(final_results, results_extra)

#saveRDS(stroke_results, file = "J:/R projects/PhD-by-chapter/Chapter6-prospective-association/rds objects/output data/HRs baseline FI (modified) and stroke by subgroups.rds")
```

```{r, fig.height=10, fig.width=8}
row_labels = final_results[, c("subheading", "label", "variable")]
row_labels$heading = "Subgroup"

forestplot1a = ckbplotr::make_forest_plot(panels        = list(stroke_results[[1]]),
                           col.key       = "variable",
                           row.labels    = row_labels,
                           row.labels.levels = c("heading", "subheading", "label"),
                           rows          = "Subgroup",
                           exponentiate  = TRUE,
                           panel.names = c("Stroke"),
                           ci.delim      = "-",
                           xlim          = c(1, 1.22),
                           xticks        = c(1, 1.1, 1.2),
                           blankrows     = c(1, 1, 0, 1),
                           scalepoints   = TRUE,
                           pointsize     = 3,
                           col.left      = c("n"),
                           col.left.heading = c("No. of\nevents"),
                           col.right.heading = c("HR (95% CI)\nper 1SD FI"),
                           col.heading.space = 1.5,
                           addtext       = list(stroke_results[[2]]),
                           right.space   = unit(35, "mm"))

forestplot2a = ckbplotr::make_forest_plot(panels        = list(ihd_results[[1]]),
                           col.key       = "variable",
                           row.labels    = row_labels,
                           row.labels.levels = c("heading", "subheading", "label"),
                           rows          = "Subgroup",
                           exponentiate  = TRUE,
                           panel.names = c("IHD"),
                           ci.delim      = "-",
                           xlim          = c(1, 1.22),
                           xticks        = c(1, 1.1, 1.2),
                           blankrows     = c(1, 1, 0, 1),
                           scalepoints   = TRUE,
                           pointsize     = 3,
                           col.left      = c("n"),
                           col.left.heading = c("No. of\nevents"),
                           col.right.heading = c("HR (95% CI)\nper 1SD FI"),
                           col.heading.space = 1.5,
                           addtext       = list(ihd_results[[2]]),
                           right.space   = unit(35, "mm")) 
```




Merge and save plot

```{r, fig.height=5, fig.width=7}
# Step 2: Remove the axis text for all but the first plot
p1 <- forestplot1a$plot

p2 <- forestplot2a$plot + 
  theme(axis.text.y = element_blank())

# Step 3: Arrange the plots using gridExtra (there may be other packages that also work). Adjust widths until the plots are the desired widths in your final output file
final_plot = gridExtra::grid.arrange(p1, p2, nrow = 1, widths = c(1, 0.7))

cowplot::save_plot("J:/R projects/PhD-by-chapter/Chapter6-prospective-association/outputs/age-at-risk stratified HRs/modified FI/HRplots_subgroups_FImodified_1SD_IHDStroke_fulladj_fatmass.png", final_plot, base_height = 5, base_aspect_ratio = 2) # aspect ratio = width/height
```
