---
title: "Age-at-risk stratified Cox PH analysis plots"
output: html_document
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Libraries

```{r}
library(ckbplotr)
library(cowplot)
library(tidyverse)
```

## Open HRs 

HRs between baseline frailty and outcomes were stored in an RDS object (nested list structure)

```{r}
#hrlist = readRDS(file = "J:/R projects/PhD-by-chapter/Chapter6-prospective-association/rds objects/output data/HRs baseline FI and outcomes.rds")
hrlist = readRDS(file = "J:/R projects/PhD-by-chapter/Chapter6-prospective-association/rds objects/output data/HRs baseline FI (modified) and stroke and IHD.rds")
hrlist = readRDS(file = "J:/R projects/PhD-by-chapter/Chapter6-prospective-association/rds objects/output data/HRs baseline SRH and stroke and IHD.rds")
```

Nested within each of the `r length(hrlist)` list elements:

```{r}
names(hrlist)
```

Are HRs between baseline frailty and the following 23 outcomes:

```{r}
names(hrlist[[1]])
```

Select IHD total and stroke total using continuous exposure

```{r}
#hrlist = hrlist[[1]][c("IHD total", "Stroke total")]
```

Rename list

```{r}
names(hrlist) = c("IHD", "Stroke")
```

## Age-at-risk stratified HRs between baseline frailty and outcomes

Cox proportional hazards models were adjusted for age-at-risk (5 year groups), sex (when appropriate), region, income, marital status, occupation, education level, hypertension diagnosis, diabetes diagnosis, alcohol category, smoking category, poor adiposity, poor diet, and physical activity.

### HRs for total participants 

Baseline exposure (SRH/FI quintiles) and outcomes

```{r eval=FALSE, fig.width=10, fig.height=10}
plots = list()

for (i in 1:length(hrlist)){
  
  title = names(hrlist[i]) 
  
  hrlist[[i]]$rf = paste0("Quintile ", c(1:5))
  
  plots[[i]] = ckbplotr::make_shape_plot(hrlist[[i]],
                            col.x          = "rf",
                            col.estimate   = "est",
                            col.stderr     = "se",
                            col.n          = "n",
                            exponentiate   = TRUE, 
                            xlims          = c(1, 5),
                            ylims          = c(0.85, 2),
                            ybreaks        = c(1, 1.5, 2),
                            xlab           = "Frailty index",
                            ylab           = "Hazard ratio (95% CI)",
                            scalepoints    = TRUE,
                            title          = title,
                            printplot      = FALSE, 
                            showcode = TRUE)
  
}

plot1 = cowplot::plot_grid(plots[[1]]$plot, plots[[2]]$plot,
                            ncol=2, nrow=2, align = "hv")

plot1

#cowplot::save_plot(paste0("J:/R projects/PhD-by-chapter/Chapter6-prospective-association/outputs/age-at-risk stratified HRs/", fi_type, " FI/HRplots_FI", fi_type, "_quintile_CVD1_fulladj.png"), plot1a, base_height = 8, base_aspect_ratio = 1)
```

*Frailty and CVD*

Plot for IHD

```{r}
# Create the ggplot
plot <- ggplot(data = hrlist[[1]][[1]],
               aes(x = rf, y = exp(est))) +

  # Plot the point estimates
  geom_point(aes(size = 1/se),
             shape = 22,
             colour = "black",
             fill = "black",
             stroke = 0.5) +
  
  # Plot point estimates text
  geom_text(aes(y = exp(est+1.96*se),
                label = format(round(exp(est), 2), nsmall = 2)),
            vjust = -0.8,
            size  = 3) +
  
  # Plot n events text
  geom_text(aes(y = exp(est-1.96*se),
                label = n),
            vjust = 1.8,
            size  = 3) +
  
  # Plot the CIs
  geom_linerange(aes(ymin = exp(est-1.96*se),
                     ymax = exp(est+1.96*se)),
                 colour = "black",
                 lwd = 0.5) +
  
  # Set the scale for the size of boxes
  scale_radius(guide  = "none",
               limits = c(0, NA_real_),
               range  = c(0, 3)) +
  
  # Use identity for aesthetic scales
  scale_shape_identity() +
  scale_colour_identity() +
  scale_fill_identity() +
  
  # Set the y-axis scale
  scale_y_continuous(trans = "log", breaks = c(1, 1.5, 2)) +
  
  # Add titles
  xlab("Frailty index quintiles") +
  ylab("Hazard ratio (95% CI)") +
  ggtitle("IHD") +

  # Plot n events text and HRs per 1SD increase in FI
  geom_text(aes(y = 1.99,
                x = 3,
                label = "(49,937 cases)"),
            size  = 3) +
  geom_text(aes(y = 1.75,
                x = 3,
                label = "HR (95%CI): 1.16 (1.14-1.17)\nper 1SD increase"),
            size  = 3.5)
  
# Plot like a CKB plot
plot_ihd = ckbplotr::plot_like_ckb(plot           = plot,
                                   xlims          = c(1, 5),
                                   ylims          = c(0.85, 2),
                                   gap            = c(0.05, 0.025),
                                   ext            = c(0.025, 0.025),
                                   ratio          = 1.5,
                                   base_size      = 11,
                                   base_line_size = 0.5)
```

Plot for stroke
```{r}
# Create the ggplot
plot <- ggplot(data = hrlist[[2]][[1]],
               aes(x = rf, y = exp(est))) +

  # Plot the point estimates
  geom_point(aes(size = 1/se),
             shape = 22,
             colour = "black",
             fill = "black",
             stroke = 0.5) +
  
  # Plot point estimates text
  geom_text(aes(y = exp(est+1.96*se),
                label = format(round(exp(est), 2), nsmall = 2)),
            vjust = -0.8,
            size  = 3) +
  
  # Plot n events text
  geom_text(aes(y = exp(est-1.96*se),
                label = n),
            vjust = 1.8,
            size  = 3) +
  
  # Plot the CIs
  geom_linerange(aes(ymin = exp(est-1.96*se),
                     ymax = exp(est+1.96*se)),
                 colour = "black",
                 lwd = 0.5) +
  
  # Set the scale for the size of boxes
  scale_radius(guide  = "none",
               limits = c(0, NA_real_),
               range  = c(0, 3)) +
  
  # Use identity for aesthetic scales
  scale_shape_identity() +
  scale_colour_identity() +
  scale_fill_identity() +
  
  # Set the y-axis scale
  scale_y_continuous(trans = "log", breaks = c(1, 1.5, 2)) +
  
  # Add titles
  xlab("Frailty index quintiles") +
  ylab("Hazard ratio (95% CI)") +
  ggtitle("Stroke") +

  # Plot n events text and HRs per 1SD increase in FI
  geom_text(aes(y = 1.99,
                x = 3,
                label = "(54,193 cases)"),
            size  = 3) +
  geom_text(aes(y = 1.75,
                x = 3,
                label = "HR (95%CI): 1.07 (1.06-1.08)\nper 1SD increase"),
            size  = 3.5)
  
  # Plot like a CKB plot
  plot_stroke = ckbplotr::plot_like_ckb(plot           = plot,
                                       xlims          = c(1, 5),
                                       ylims          = c(0.85, 2),
                                       gap            = c(0.05, 0.025),
                                       ext            = c(0.025, 0.025),
                                       ratio          = 1.5,
                                       base_size      = 11,
                                       base_line_size = 0.5)
 
```

Combine and save plot
```{r, fig.width=10, fig.height=10}
p = cowplot::plot_grid(plot_stroke, plot_ihd,
                           ncol=2, nrow=1, align = "hv")

title = ggdraw() + draw_label("Figure 1. Association between baseline frailty index, stroke, and IHD", fontface='bold')

plot1 = plot_grid(title, p, ncol=1, rel_heights=c(0.1, 1)) # rel_heights values control title margins
plot1

cowplot::save_plot("J:/R projects/PhD-by-chapter/Chapter6-prospective-association/outputs/age-at-risk stratified HRs/modified FI/HRplots_FImodified_quintile_IHDStroke_fulladj.png", plot1, base_height = 6, base_aspect_ratio = 1.5)
```

*SRH and CVD*

Plot for IHD

```{r}
# Create the ggplot
plot <- ggplot(data = hrlist[[1]][[1]],
               aes(x = rf, y = exp(est))) +

  # Plot the point estimates
  geom_point(aes(size = 1/se),
             shape = 22,
             colour = "black",
             fill = "black",
             stroke = 0.5) +
  
  # Plot point estimates text
  geom_text(aes(y = exp(est+1.96*se),
                label = format(round(exp(est), 2), nsmall = 2)),
            vjust = -0.8,
            size  = 3) +
  
  # Plot n events text
  geom_text(aes(y = exp(est-1.96*se),
                label = n),
            vjust = 1.8,
            size  = 3) +
  
  # Plot the CIs
  geom_linerange(aes(ymin = exp(est-1.96*se),
                     ymax = exp(est+1.96*se)),
                 colour = "black",
                 lwd = 0.5) +
  
  # Set the scale for the size of boxes
  scale_radius(guide  = "none",
               limits = c(0, NA_real_),
               range  = c(0, 3)) +
  
  # Use identity for aesthetic scales
  scale_shape_identity() +
  scale_colour_identity() +
  scale_fill_identity() +
  
  # Set the y-axis scale
  scale_y_continuous(trans = "log", breaks = c(1, 1.5, 2)) +
  # Set the x-axis scale
  scale_x_discrete(labels = c("Excellent", "Good", "Fair", "Poor")) +
  
  # Add titles
  xlab("Self-rated health") +
  ylab("Hazard ratio (95% CI)") +
  ggtitle("IHD") +

  # Plot n events text and HRs per 1SD increase in FI
  geom_text(aes(y = 1.99,
                x = 2.5,
                label = "(49,937 cases)"),
            size  = 3) +
  geom_text(aes(y = 1.75,
                x = 2.5,
                label = "HR (95%CI): 1.15 (1.14-1.16)\nper 1SD increase"),
            size  = 3.5)
  
# Plot like a CKB plot
plot_ihd = ckbplotr::plot_like_ckb(plot           = plot,
                                   xlims          = c(1, 4),
                                   ylims          = c(0.85, 2),
                                   gap            = c(0.05, 0.025),
                                   ext            = c(0.025, 0.025),
                                   ratio          = 1.5,
                                   base_size      = 11,
                                   base_line_size = 0.5)
```

Plot for stroke
```{r}
# Create the ggplot
plot <- ggplot(data = hrlist[[2]][[1]],
               aes(x = rf, y = exp(est))) +

  # Plot the point estimates
  geom_point(aes(size = 1/se),
             shape = 22,
             colour = "black",
             fill = "black",
             stroke = 0.5) +
  
  # Plot point estimates text
  geom_text(aes(y = exp(est+1.96*se),
                label = format(round(exp(est), 2), nsmall = 2)),
            vjust = -0.8,
            size  = 3) +
  
  # Plot n events text
  geom_text(aes(y = exp(est-1.96*se),
                label = n),
            vjust = 1.8,
            size  = 3) +
  
  # Plot the CIs
  geom_linerange(aes(ymin = exp(est-1.96*se),
                     ymax = exp(est+1.96*se)),
                 colour = "black",
                 lwd = 0.5) +
  
  # Set the scale for the size of boxes
  scale_radius(guide  = "none",
               limits = c(0, NA_real_),
               range  = c(0, 3)) +
  
  # Use identity for aesthetic scales
  scale_shape_identity() +
  scale_colour_identity() +
  scale_fill_identity() +
  
  # Set the y-axis scale
  scale_y_continuous(trans = "log", breaks = c(1, 1.5, 2)) +
  # Set the x-axis scale
  scale_x_discrete(labels = c("Excellent", "Good", "Fair", "Poor")) +
  
  # Add titles
  xlab("Self-rated health") +
  ylab("Hazard ratio (95% CI)") +
  ggtitle("Stroke") +

  # Plot n events text and HRs per 1SD increase in FI
  geom_text(aes(y = 1.99,
                x = 2.5,
                label = "(54,193 cases)"),
            size  = 3) +
  geom_text(aes(y = 1.75,
                x = 2.5,
                label = "HR (95%CI): 1.07 (1.06-1.08)\nper 1SD increase"),
            size  = 3.5)
  
  # Plot like a CKB plot
  plot_stroke = ckbplotr::plot_like_ckb(plot           = plot,
                                       xlims          = c(1, 4),
                                       ylims          = c(0.85, 2),
                                       gap            = c(0.05, 0.025),
                                       ext            = c(0.025, 0.025),
                                       ratio          = 1.5,
                                       base_size      = 11,
                                       base_line_size = 0.5)
 
```

Combine and save plot
```{r, fig.width=10, fig.height=10}
p = cowplot::plot_grid(plot_stroke, plot_ihd,
                           ncol=2, nrow=1, align = "hv")

title = ggdraw() + draw_label("Figure 1. Association between baseline self-rated health, stroke, and IHD", fontface='bold')

plot1 = plot_grid(title, p, ncol=1, rel_heights=c(0.1, 1)) # rel_heights values control title margins
plot1

cowplot::save_plot("J:/R projects/PhD-by-chapter/Chapter6-prospective-association/outputs/age-at-risk stratified HRs//HRplots_SRH_IHDStroke_fulladj.png", plot1, base_height = 6, base_aspect_ratio = 1.5)
```