---
title: "Flow diagram of participants"
output: html_document
date: '2022-03-14'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Required libraries
```{r}
library(kableExtra)
library(knitr)
library(tidyverse)
```

## Open data

### Baseline data

```{r}
# Define pathway of raw dataset
mypath = "J:/R projects/PhD-by-chapter/data/raw_data/"

data_base = read.csv(file = paste0(mypath, 'data_baseline_questionnaires.csv'))
```

### Definitions

```{r}
# Column details
column_det = read.csv(file = paste0(mypath, 'column_details.csv'))

# Value definitions
value_def = read.csv(file = paste0(mypath, 'value_definitions.csv'))
```

### Select variables of interest

```{r}
varnames = c("csid",
  "age_at_study_date_x100",
  "region_code",
  "is_female",
  
  # demographics
  "household_income",
  "marital_status",
  "occupation",
  "highest_education",
  
  # disease
  "chd_diag",
  "stroke_or_tia_diag",
  
  # covariates
  "bmi_calc", "waist_hip_ratio",
  "met",
  "sbp_mean", "dbp_mean", "hypertension_diag", "used_blood_pressure_drugs",
  "fasting_glucose_x10", "random_glucose_x10", "diabetes_diag", "taking_insulin",
  "smoking_category", "smoking_stopped_reason",
  "alcohol_category",
  "diet_freq_fresh_fruit", "diet_freq_fresh_veg", "diet_freq_meat",
  
  # frailty-related
  "emph_bronc_diag", 
  "tb_diag", 
  "asthma_diag", 
  "peptic_ulcer_diag", 
  "gall_diag", 
  "rheum_arthritis_diag", 
  "fracture_diag", 
  "neurasthenia_diag", 
  "cancer_diag", 
  "kidney_dis_diag", 
  "continuous_pain", 
  #"chd_diag", 
  #"stroke_or_tia_diag", 
  "bowel_movement_freq", 
  "heart_rate_mean", 
  #"bmi", 
  "cough_freq", 
  "gum_bleed_freq", 
  #"diabetes", 
  #"hypertension", 
  #"met_low", 
  "self_rated_health", 
  "has_copd", 
  "sleep_affects_day", "sleep_delayed_fitful", "sleep_waking_too_early", 
  #"waist_to_hip",
  "walking_short_of_breath", "walking_slow_down", 
  "tried_to_reduce_weight", "weight_changed"
)

mydata = data_base[, varnames]
```

## Count participants for flow diagram

Total participants

```{r}
nrow(mydata)
```

## Missing data

```{r}
# Count number of missing values per variable
na_count = sapply(mydata, function(x) sum(length(which(is.na(x)))))

na_count = data.frame(na_count)

kable(na_count)
```

### Exclusion by baseline disease 

Has CHD at baseline
```{r}
table(mydata$chd_diag)
```
Has stroke/TIA at baseline
```{r}
table(mydata$stroke_or_tia_diag)
```
Has CHD or stroke/TIA at baseline
```{r}
mydata$chd_stroke_tia = if_else(mydata$chd_diag==1 | mydata$stroke_or_tia_diag==1, 1, 0)
table(mydata$chd_stroke_tia)
```
CVD-free participants at baseline
```{r}
mydata2 = mydata[mydata$chd_stroke_tia==0,]
nrow(mydata2)
```

### Exclusion by baseline variables

```{r}
mydata_num = mydata2[, c("age_at_study_date_x100", "bmi_calc", "met", "sbp_mean", "dbp_mean", "fasting_glucose_x10", "random_glucose_x10", "heart_rate_mean")]


sapply(names(mydata_num), 
       function(x) {
         hist(mydata_num[[x]], main = x)
       }
)
```

Blood pressure SBP<80 or ≥250 or DBP<40 or ≥150 (Lacey et al., 2018)

```{r}
length(which(mydata2$sbp_mean<80))
length(which(mydata2$sbp_mean>=250))
length(which(mydata2$dbp_mean<40))
length(which(mydata2$dbp_mean>=150))

mydata2$extreme_bp = if_else(mydata2$sbp_mean<80 | mydata2$sbp_mean>=250 | mydata2$dbp_mean<40 | mydata2$dbp_mean>=150, 1, 0)

length(which(mydata2$extreme_bp==1))
```

## Participants included for main analysis

Excluded the following:

  - 2 missing BMI
  - 23129 has CHD or stroke/TIA at baseline
  - 147 extreme BP values
  
```{r}
ptlist_main_analysis = mydata2[mydata2$extreme_bp==0 & !is.na(mydata2$bmi_calc), ] %>% pull(csid) 

length(ptlist_main_analysis)
```
Save list of participants included for main analysis

```{r}
saveRDS(ptlist_main_analysis, file = "J:/R projects/PhD-by-chapter/rds objects/ptlist_main_analysis.rds")
```

