---
title: "Table 1"
output: 
  html_document:
    number_sections: TRUE
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE) # echo=FALSE hides codes
```

Aim of the markdown is to create various 'Table 1's for CKB data chapter of my thesis

```{r packages, include=FALSE}
# Download packages required for this 
library(data.table)   # for fread() to import specific columns
library(knitr)        # for kable()
library(tableone)     # for CreateTableOne()
library(tidyverse)
```

Import pre-processed data

```{r}
# Data consisting of frailty variables and frailty scores at the three different surveys in a wide format
mydata = readRDS("J:/R projects/PhD-by-chapter/rds objects/covariates_pretty_levels_for_everyone.rds")
```

## Proportion of study participants in study region by sex

```{r}
table1 = tableone::CreateTableOne(vars = "region_code", strata = "is_female", data = mydata)

table1 = print(table1, exact=c("region_code"))
```

## Baseline demographic and lifestyle characteristics of recruited participants, overall and by sex

```{r}
myvars = c("age_at_study_date", "age_group", "region_is_urban", "marital_status_group", "highest_education", "household_income", "occupation", 
           "smoking_group", "alcohol_category",
           "daily_fresh_fruit", "daily_fresh_veg", "daily_meat",
           "met", "self_rated_health",
           "asthma_diag", "cancer_diag", "chd_diag",  "cirrhosis_hep_diag", "copd_diag", "diabetes_diag", "emph_bronc_diag", "fracture_diag", "gall_diag",  "hypertension_diag", "kidney_dis_diag", "neurasthenia_diag", "peptic_ulcer_diag", "rheum_arthritis_diag", "rheum_heart_dis_diag", "stroke_or_tia_diag", "tb_diag")

table2_sex = tableone::CreateTableOne(vars = myvars, strata = "is_female", data = mydata)
table2_tot = tableone::CreateTableOne(vars = myvars, data = mydata)

table2_sex = print(table2_sex, nonnormal = c("met"))
table2_tot = print(table2_tot, nonnormal = c("met"))
```

## Distribution of physical measurements, overall and by sex

```{r}
myvars = c("bmi_calc", "bmi_group", "waist_hip_ratio", "whr_group", "fat_percent", "fat_body_mass_kg", "lean_body_mass_kg", 
           "sbp_mean", "sbp_group", "random_glucose", "random_glucose_group")

table3_sex = tableone::CreateTableOne(vars = myvars, strata = "is_female", data = mydata)
table3_tot = tableone::CreateTableOne(vars = myvars, data = mydata)

table3_sex = print(table3_sex)
table3_tot = print(table3_tot)
```

## Baseline demographic and lifestyle characteristics of recruited participants, by first resurvey follow-up

```{r}
myvars = c("age_at_study_date", "age_group", "region_is_urban", "marital_status_group", "highest_education", "household_income", "occupation", 
           "smoking_group", "alcohol_category",
           "daily_fresh_fruit", "daily_fresh_veg", "daily_meat",
           "met", "self_rated_health",
           "asthma_diag", "cancer_diag", "chd_diag",  "cirrhosis_hep_diag", "copd_diag", "diabetes_diag", "emph_bronc_diag", "fracture_diag", "gall_diag",  "hypertension_diag", "kidney_dis_diag", "neurasthenia_diag", "peptic_ulcer_diag", "rheum_arthritis_diag", "rheum_heart_dis_diag", "stroke_or_tia_diag", "tb_diag")

table4_tot = tableone::CreateTableOne(vars = myvars, strata = "fu_resurvey1", data = mydata)
table4_sex = tableone::CreateTableOne(vars = myvars, strata = c("fu_resurvey1", "is_female"), data = mydata)

table4_tot = print(table4_tot, nonnormal = c("met"))
table4_sex = print(table4_sex, nonnormal = c("met"))
```

## Distribution of physical measurements, overall and by sex

```{r}
myvars = c("bmi_calc", "bmi_group", "waist_hip_ratio", "whr_group", "fat_percent", "fat_body_mass_kg", "lean_body_mass_kg", 
           "sbp_mean", "sbp_group", "random_glucose", "random_glucose_group")

table5_tot = tableone::CreateTableOne(vars = myvars, strata = "fu_resurvey1", data = mydata)
table5_sex = tableone::CreateTableOne(vars = myvars, strata = c("fu_resurvey1", "is_female"), data = mydata)

table5_tot = print(table5_tot)
table5_sex = print(table5_sex)
```

Save outputs

```{r}
write.csv(table1, "J:/R projects/PhD-by-chapter/Chapter4-CKB/outputs/Table 1s/n (pct) in each region by sex.csv")
write.csv(table2_tot, "J:/R projects/PhD-by-chapter/Chapter4-CKB/outputs/Table 1s/baseline characteristic total CKB.csv")
write.csv(table2_sex, "J:/R projects/PhD-by-chapter/Chapter4-CKB/outputs/Table 1s/baseline characteristic by sex.csv")
write.csv(table3_tot, "J:/R projects/PhD-by-chapter/Chapter4-CKB/outputs/Table 1s/baseline adiposity and sbp total CKB.csv")
write.csv(table3_sex, "J:/R projects/PhD-by-chapter/Chapter4-CKB/outputs/Table 1s/baseline adiposity and sbp by sex.csv")
write.csv(table4_tot, "J:/R projects/PhD-by-chapter/Chapter4-CKB/outputs/Table 1s/baseline characteristic by resurvey1 follow-up.csv")
write.csv(table4_sex, "J:/R projects/PhD-by-chapter/Chapter4-CKB/outputs/Table 1s/baseline characteristic by resurvey1 follow-up and sex.csv")
write.csv(table5_tot, "J:/R projects/PhD-by-chapter/Chapter4-CKB/outputs/Table 1s/baseline adiposity and sbp by resurvey1 follow-up.csv")
write.csv(table5_sex, "J:/R projects/PhD-by-chapter/Chapter4-CKB/outputs/Table 1s/baseline adiposity and sbp by resurvey1 follow-up and sex.csv")
```
