---
title: "Variability in frailty"
output: html_document
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Libraries

```{r}
library(data.table)     # for fread()
library(knitr)
library(tidyverse)
```

## Data

### Open data

```{r}
ptlist_main_analysis = readRDS("J:/R projects/PhD-by-chapter/rds objects/ptlist_main_analysis.rds") 
widedata = readRDS("J:/R projects/PhD-by-chapter/data/processed_data/data_wide_fi_no_cv.rds")

# Study entry dates
study_entry_base = data.table::fread("J:/R projects/PhD-by-chapter/data/raw_data/data_baseline_questionnaires.csv", select = c("csid", "study_date"))
study_entry_resurvey1 = data.table::fread("J:/R projects/PhD-by-chapter/data/raw_data/data_resurvey1_questionnaires.csv", select = c("csid", "study_date"))
study_entry_resurvey2 = data.table::fread("J:/R projects/PhD-by-chapter/data/raw_data/data_resurvey2_questionnaires.csv", select = c("csid", "study_date"))

widedata = widedata %>% 
  left_join(select(study_entry_base, csid, study_date_base = study_date), by = "csid") %>% 
  left_join(select(study_entry_resurvey1, csid, study_date_resurvey1 = study_date), by = "csid") %>% 
  left_join(select(study_entry_resurvey2, csid, study_date_resurvey2 = study_date), by = "csid") %>% 
  mutate(across(contains("_date"), ~as.Date(.)))
  
widedata = widedata[widedata$csid %in% ptlist_main_analysis, ]
```

### Process data

Add new variable (FI quintile groups at baseline)

```{r}
widedata$fi5= factor(ntile(widedata$b_fi.score, 5)) # split FI into quintile groups based on baseline scores
```

Calculate number of days between each study entry

```{r}
widedata$days_from_base_to_resurvey1 =  widedata$study_date_resurvey1 - widedata$study_date_base
widedata$days_from_base_to_resurvey2 =  widedata$study_date_resurvey2 - widedata$study_date_base
widedata$days_from_resurvey1_to_resurvey2 =  widedata$study_date_resurvey2 - widedata$study_date_resurvey1
```

### Subset data 

1. Has baseline and 1st resurvey data

```{r}
anyNA(widedata$b_fi.score)

temp = pull(widedata[which(!is.na(widedata$r1_fi.score)), ], csid)
widedata1 = widedata[widedata$csid %in% temp, c("csid", "b_fi.score", "r1_fi.score", "fi5", "days_from_base_to_resurvey1")]
```

`r nrow(widedata1)` participants

2. Has baseline and 2nd resurvey data

```{r}
temp = pull(widedata[which(!is.na(widedata$r2_fi.score)), ], csid)

widedata2 = widedata[widedata$csid %in% temp, c("csid", "b_fi.score", "r2_fi.score", "fi5", "days_from_base_to_resurvey2")]
```

`r nrow(widedata2)` participants

2. Has baseline and 1st and 2nd resurvey data

```{r}
temp = pull(widedata[which(!(is.na(widedata$r2_fi.score) | is.na(widedata$r1_fi.score))), ], csid)

widedata3 = widedata[widedata$csid %in% temp, c("csid", "b_fi.score", "r1_fi.score", "r2_fi.score", "fi5", "days_from_base_to_resurvey2", "days_from_base_to_resurvey1")]
```

`r nrow(widedata3)` participants

## Assess variability in frailty index scores

### Change in mean FI between baseline and first resurvey in quintile groups defined by baseline frailty

```{r}
# Make data long
fivar = widedata1 %>%
  pivot_longer(cols = contains("fi.score"), names_to = "survey", values_to = "fi.score") %>% 
  mutate(survey = if_else(survey == "b_fi.score", "Baseline", "First resurvey"))

# Compute mean FI score and interval between baseline and resurvey by baseline FI quintile groups
to_plot = fivar %>% 
  group_by(survey, fi5) %>% 
  summarise(mean = mean(fi.score),
            mean_interval_days = mean(days_from_base_to_resurvey1))

# Make intervals start from 0 at baseline
to_plot = to_plot %>% 
  mutate(mean_interval_days = if_else(survey == "Baseline", lubridate::make_difftime(0, units = "day"), mean_interval_days))

p = to_plot %>% 
  ggplot(aes(x = mean_interval_days/365.25, y = mean, color = survey, label = round(mean,2))) +
  geom_point(aes(fill = survey), shape = 22) +
  geom_line(aes(group = fi5), color = "black") +
  geom_text(size = 3, color = "black", vjust = 0, nudge_y = 0.005) +
  scale_x_continuous(breaks = c(0, 0.5, 1, 1.5, 2, 2.5, 3)) +
  labs(x = "Mean interval between measurements (years)",
       y = "Mean frailty index score") +
  theme_classic() +
  theme(aspect.ratio = 4/3) # ratio y/x
p
```

### Change in mean FI between baseline and second resurvey in quintile groups defined by baseline frailty

```{r}
# Make data long
fivar = widedata2 %>%
  pivot_longer(cols = contains("fi.score"), names_to = "survey", values_to = "fi.score") %>% 
  mutate(survey = if_else(survey == "b_fi.score", "Baseline", "Second resurvey"))

# Compute mean FI score and interval between baseline and resurvey by baseline FI quintile groups
to_plot = fivar %>% 
  group_by(survey, fi5) %>% 
  summarise(mean = mean(fi.score),
            mean_interval_days = mean(days_from_base_to_resurvey2))

# Make intervals start from 0 at baseline
to_plot = to_plot %>% 
  mutate(mean_interval_days = if_else(survey == "Baseline", lubridate::make_difftime(0, units = "day"), mean_interval_days))

p = to_plot %>% 
  ggplot(aes(x = mean_interval_days/365.25, y = mean, color = survey, label = round(mean,2))) +
  geom_point(aes(fill = survey), shape = 22) +
  geom_line(aes(group = fi5), color = "black") +
  geom_text(size = 3, color = "black", vjust = 0, nudge_y = 0.005) +
  scale_x_continuous(breaks = c(0, 1, 2, 3, 4, 5, 6, 7, 8)) +
  labs(x = "Mean interval between measurements (years)",
       y = "Mean frailty index score") +
  theme_classic() +
  theme(aspect.ratio = 4/3) # ratio y/x
p
```

### Change in mean FI between baseline and first and second resurveys in quintile groups defined by baseline frailty

```{r}
# Make data long, compute mean FI score by survey and quintile group
temp = widedata3 %>%  
  select(-contains("days")) %>% 
  pivot_longer(cols = contains("fi.score"), names_to = "survey", values_to = "fi.score") %>% 
  mutate(survey = case_when(survey == "b_fi.score" ~ "Baseline", 
                            survey == "r1_fi.score" ~ "First resurvey",
                            TRUE ~ "Second resurvey")) %>% 
  group_by(survey, fi5) %>% 
  summarise(mean = mean(fi.score))

# Make data long, compute mean interval by survey and quintile group
temp1 = widedata3 %>%
  select(-contains("fi.score")) %>% 
  pivot_longer(cols = contains("days"), names_to = "interval", values_to = "days") %>% 
  mutate(survey = if_else(interval=="days_from_base_to_resurvey1", "First resurvey", "Second resurvey")) %>% 
  group_by(survey, fi5) %>% 
  summarise(mean_interval_days = mean(days))

# Merge data
to_plot = temp %>% 
  left_join(temp1, by = c("survey", "fi5")) %>% 
  mutate(mean_interval_days = if_else(survey == "Baseline", lubridate::make_difftime(0, units = "day"), mean_interval_days)) # Make intervals start from 0 at baseline

p = to_plot %>% 
  ggplot(aes(x = mean_interval_days/365.25, y = mean, color = survey, label = round(mean,2))) +
  geom_point(aes(fill = survey), shape = 22) +
  geom_line(aes(group = fi5), color = "black") +
  geom_text(size = 3, color = "black", vjust = 0, nudge_y = 0.005) +
  scale_x_continuous(breaks = c(0, 1, 2, 3, 4, 5, 6, 7, 8)) +
  labs(x = "Mean interval between measurements (years)",
       y = "Mean frailty index score") +
  theme_classic() +
  theme(aspect.ratio = 4/3) # ratio y/x
p
```