---
title: "Variability in frailty"
output: html_document
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Libraries

```{r}
library(data.table)     # for fread()
library(knitr)
library(tidyverse)
```

## Data

### Open data

```{r}
#ptlist_main_analysis = readRDS("J:/R projects/PhD-by-chapter/rds objects/ptlist_main_analysis.rds") 
fi.nocv = readRDS("J:/R projects/PhD-by-chapter/data/scripts/data_wide_fi_no_cv.rds")
fi.noad = readRDS("J:/R projects/PhD-by-chapter/data/scripts/data_wide_fi_no_adiposity.rds")
fi.no21 = readRDS("J:/R projects/PhD-by-chapter/data/scripts/data_wide_fi_no_21.rds")
fi.full = readRDS("J:/R projects/PhD-by-chapter/data/scripts/data_wide_fi_full.rds")

# Study entry dates
study_entry_base = data.table::fread("J:/R projects/PhD-by-chapter/data/raw_data/data_baseline_questionnaires.csv", select = c("csid", "study_date"))
study_entry_resurvey1 = data.table::fread("J:/R projects/PhD-by-chapter/data/raw_data/data_resurvey1_questionnaires.csv", select = c("csid", "study_date"))
study_entry_resurvey2 = data.table::fread("J:/R projects/PhD-by-chapter/data/raw_data/data_resurvey2_questionnaires.csv", select = c("csid", "study_date"))
```

### Process data

```{r}
process_data = function(data){
  
  # Add dates
  data = data %>% 
   left_join(select(study_entry_base, csid, study_date_base = study_date), by = "csid") %>% 
   left_join(select(study_entry_resurvey1, csid, study_date_resurvey1 = study_date), by = "csid") %>% 
   left_join(select(study_entry_resurvey2, csid, study_date_resurvey2 = study_date), by = "csid") %>% 
   mutate(across(contains("_date"), ~as.Date(.)))
  
  # Calculate number of days between each study entry
  data$days_from_base_to_resurvey1 =  data$study_date_resurvey1 - data$study_date_base
  data$days_from_base_to_resurvey2 =  data$study_date_resurvey2 - data$study_date_base
  data$days_from_resurvey1_to_resurvey2 =  data$study_date_resurvey2 - data$study_date_resurvey1
  
  # Add new variable (FI quintile groups at baseline
  data$fi5= factor(ntile(data$b_fi.score, 5)) # split FI into quintile groups based on baseline scores

  return(data)
}

fi.full = process_data(fi.full)
fi.nocv = process_data(fi.nocv)
fi.noad = process_data(fi.noad)
fi.no21 = process_data(fi.no21)
```

### Subset data 

1. Has baseline and 1st resurvey data

```{r}
subset_data_base_and_1st = function(widedata){
  
  temp = pull(widedata[which(!is.na(widedata$r1_fi.score)), ], csid)
  output = widedata[widedata$csid %in% temp, c("csid", "b_fi.score", "r1_fi.score", "fi5", "days_from_base_to_resurvey1")]
  
  return(output)
}

widedata_br1_fi = subset_data_base_and_1st(fi.full)
widedata_br1_mfi1 = subset_data_base_and_1st(fi.nocv)
widedata_br1_mfi2 = subset_data_base_and_1st(fi.noad)
widedata_br1_mfi3 = subset_data_base_and_1st(fi.no21)

nrow(widedata_br1_fi)
```

2. Has baseline and 2nd resurvey data

```{r}
subset_data_base_and_2nd = function(widedata){
  
  temp = pull(widedata[which(!is.na(widedata$r2_fi.score)), ], csid)
  output = widedata[widedata$csid %in% temp, c("csid", "b_fi.score", "r2_fi.score", "fi5", "days_from_base_to_resurvey2")]
  
  return(output)
}

widedata_br2_fi = subset_data_base_and_2nd(fi.full)
widedata_br2_mfi1 = subset_data_base_and_2nd(fi.nocv)
widedata_br2_mfi2 = subset_data_base_and_2nd(fi.noad)
widedata_br2_mfi3 = subset_data_base_and_2nd(fi.no21)

nrow(widedata_br2_fi)
```

2. Has baseline and 1st and 2nd resurvey data

```{r}
subset_data_1st_and_2nd = function(widedata){
  
  temp = pull(widedata[which(!(is.na(widedata$r2_fi.score) | is.na(widedata$r1_fi.score))), ], csid)
  output = widedata[widedata$csid %in% temp, c("csid", "b_fi.score", "r1_fi.score", "r2_fi.score", "fi5", "days_from_base_to_resurvey2", "days_from_base_to_resurvey1")]
  
  return(output)
}

widedata_r1r2_fi = subset_data_1st_and_2nd(fi.full)
widedata_r1r2_mfi1 = subset_data_1st_and_2nd(fi.nocv)
widedata_r1r2_mfi2 = subset_data_1st_and_2nd(fi.noad)
widedata_r1r2_mfi3 = subset_data_1st_and_2nd(fi.no21)

nrow(widedata_r1r2_fi)
```

## Assess variability in frailty index scores

### Change in mean FI between baseline and first resurvey in quintile groups defined by baseline frailty

```{r}
mean_frailty_change = function(data, title, string_wave1, string_wave2, days_n, x_breaks, ylab){
  
  long_data = data %>% 
    pivot_longer(cols = contains("fi.score"), names_to = "survey", values_to = "fi.score") %>% 
    mutate(survey = if_else(survey == "b_fi.score", string_wave1, string_wave2))
  
  # Compute mean FI score and interval between baseline and resurvey by baseline FI quintile groups
  to_plot = long_data %>% 
    group_by(survey, fi5) %>% 
    summarise(mean = mean(fi.score),
              mean_interval_days = mean({{days_n}}))
  
  # Make intervals start from 0 at baseline
  to_plot = to_plot %>% 
    mutate(mean_interval_days = if_else(survey == "Baseline", lubridate::make_difftime(0, units = "day"), mean_interval_days))

  p = to_plot %>% 
    ggplot(aes(x = mean_interval_days/365.25, y = mean, color = survey, label = round(mean,2))) +
    geom_point(aes(fill = survey), shape = 22) +
    geom_line(aes(group = fi5), color = "black") +
    geom_text(size = 3, color = "black", vjust = 0, nudge_y = 0.005) +
    scale_y_continuous(limits = c(0, 0.205), breaks = c(0, 0.05, 0.10, 0.15, 0.20)) +
    scale_x_continuous(breaks = x_breaks) +
    labs(x = "Mean interval between measurements (years)",
         y = ylab) +
    ggtitle(title) +
    theme_classic() +
    theme(aspect.ratio = 4/3, # ratio y/x
          legend.position = "none",
          plot.title = element_text(hjust=0.5),
          axis.title.x = element_text(size=9)) 
  
  print(p)
  
  output = list(to_plot, p)
  
  return(output)    
  
}

mean_frailty_change_br1_fi = mean_frailty_change(widedata_br1_fi, "Original FI", "Baseline", "1st resurvey", days_from_base_to_resurvey1, c(0, 0.5, 1, 1.5, 2, 2.5, 3), "Mean FI score")
mean_frailty_change_br1_mfi1 = mean_frailty_change(widedata_br1_mfi1, "Modified FI-1", "Baseline", "1st resurvey", days_from_base_to_resurvey1, c(0, 0.5, 1, 1.5, 2, 2.5, 3), "Mean FI score")
mean_frailty_change_br1_mfi2 = mean_frailty_change(widedata_br1_mfi2, "Modified FI-2", "Baseline", "1st resurvey", days_from_base_to_resurvey1, c(0, 0.5, 1, 1.5, 2, 2.5, 3), "Mean FI score")
mean_frailty_change_br1_mfi3 = mean_frailty_change(widedata_br1_mfi3, "Modified FI-3", "Baseline", "1st resurvey", days_from_base_to_resurvey1, c(0, 0.5, 1, 1.5, 2, 2.5, 3), "Mean FI score")
```

Combine and save plots

```{r}
p = cowplot::plot_grid(mean_frailty_change_br1_fi[[2]], mean_frailty_change_br1_mfi1[[2]], 
                       mean_frailty_change_br1_mfi2[[2]], mean_frailty_change_br1_mfi3[[2]],
                       nrow = 2,
                       align = "h") +
     theme(plot.margin = unit(c(t=0,r=0.5,b=0,l=0), "cm"))

p

ggsave(filename = "plot_change_in_fi_br1.png", plot = p, path = "J:/R projects/PhD-by-chapter/Chapter5-frailtyindex/Variability", width = 6.5, height = 8)
```

### Change in mean FI between baseline and second resurvey in quintile groups defined by baseline frailty

```{r}
mean_frailty_change_br2_fi = mean_frailty_change(widedata_br2_fi, "Original FI", "Baseline", "2nd resurvey", days_from_base_to_resurvey2, seq(0,8,2), "Mean FI score")
mean_frailty_change_br2_mfi1 = mean_frailty_change(widedata_br2_mfi1, "Modified FI-1", "Baseline", "2nd resurvey", days_from_base_to_resurvey2, seq(0,8,2), "Mean FI score")
mean_frailty_change_br2_mfi2 = mean_frailty_change(widedata_br2_mfi2, "Modified FI-2", "Baseline", "2nd resurvey", days_from_base_to_resurvey2, seq(0,8,2), "Mean FI score")
mean_frailty_change_br2_mfi3 = mean_frailty_change(widedata_br2_mfi3, "Modified FI-3", "Baseline", "2nd resurvey", days_from_base_to_resurvey2, seq(0,8,2), "Mean FI score")
```
Combine and save plots

```{r}
p = cowplot::plot_grid(mean_frailty_change_br2_fi[[2]], mean_frailty_change_br2_mfi1[[2]], 
                       mean_frailty_change_br2_mfi2[[2]], mean_frailty_change_br2_mfi3[[2]],
                       nrow = 2,
                       align = "h") +
     theme(plot.margin = unit(c(t=0,r=0.5,b=0,l=0), "cm"))

p

ggsave(filename = "plot_change_in_fi_br2.png", plot = p, path = "J:/R projects/PhD-by-chapter/Chapter5-frailtyindex/Variability", width = 6.5, height = 8)
```

### Change in mean FI between baseline and first and second resurveys in quintile groups defined by baseline frailty

```{r}
# Make data long, compute mean FI score by survey and quintile group
temp = widedata3 %>%  
  select(-contains("days")) %>% 
  pivot_longer(cols = contains("fi.score"), names_to = "survey", values_to = "fi.score") %>% 
  mutate(survey = case_when(survey == "b_fi.score" ~ "Baseline", 
                            survey == "r1_fi.score" ~ "First resurvey",
                            TRUE ~ "Second resurvey")) %>% 
  group_by(survey, fi5) %>% 
  summarise(mean = mean(fi.score))

# Make data long, compute mean interval by survey and quintile group
temp1 = widedata3 %>%
  select(-contains("fi.score")) %>% 
  pivot_longer(cols = contains("days"), names_to = "interval", values_to = "days") %>% 
  mutate(survey = if_else(interval=="days_from_base_to_resurvey1", "First resurvey", "Second resurvey")) %>% 
  group_by(survey, fi5) %>% 
  summarise(mean_interval_days = mean(days))

# Merge data
to_plot = temp %>% 
  left_join(temp1, by = c("survey", "fi5")) %>% 
  mutate(mean_interval_days = if_else(survey == "Baseline", lubridate::make_difftime(0, units = "day"), mean_interval_days)) # Make intervals start from 0 at baseline

p = to_plot %>% 
  ggplot(aes(x = mean_interval_days/365.25, y = mean, color = survey, label = round(mean,2))) +
  geom_point(aes(fill = survey), shape = 22) +
  geom_line(aes(group = fi5), color = "black") +
  geom_text(size = 3, color = "black", vjust = 0, nudge_y = 0.005) +
  scale_x_continuous(breaks = c(0, 1, 2, 3, 4, 5, 6, 7, 8)) +
  labs(x = "Mean interval between measurements (years)",
       y = "Mean frailty index score") +
  theme_classic() +
  theme(aspect.ratio = 4/3) # ratio y/x
p
```