---
title: 'Cox regression between baseline frailty and outcomes: age-at-risk stratification
  (continuous FI)'
output: html_document
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Required libraries

```{r library}
library(broom)      # for tidy() to get clean HRs and confidence intervals from model fit
library(cowplot)
library(data.table)
library(Epi)        # for float(), which compute floating absolute risks using the method by Easton et al.
library(kableExtra)
library(knitr)
library(lubridate)
library(survival)   # for Surv(), cox regression
library(tidyverse)
```

## Open and process data

Run external script to open and process data

```{r open-data}
# Source script that creates working dataset that includes baseline frailty scores and CVD outcomes with key baseline covariates
source("J:/R projects/PhD-by-chapter/Chapter6-prospective-association/code/source/script to open and process dataset to analyse baseline frailty and outcomes.R")
```

Data structure

```{r data-structure}
str(mydata)
```

Number of participants included:

  - FI analysis: `r nrow(mydata)`

Further process data for age-at-risk analysis to have following variables:

  - csid
  - dob_anon
  - study_date
  - endpoint - binary variable indicating the endpoint (1/0)
  - endpoint_date - date of endpoint event or censoring
  - frailty quintile group
  - outcome_type - indicate outcome type, e.g. stroke, MI, etc.

```{r data-process, message=FALSE, warning=FALSE}
mydata = mydata %>% 
  mutate_at(vars(contains("_ind")), ~replace(., is.na(.), 0)) # replace missing values with zeros for dummy variables 

mydata = mydata %>% 
  mutate_at(vars(contains("total_date"), contains("incidence_date"), contains("fatal_date"), date_of_death), ~replace(., is.na(.), censoring_date)) # if empty dates (i.e. no event), then replace empty values with censoring date

kable(head(mydata)) %>% kableExtra::kable_styling()
```
Numbers and other summary stats
  - Total n
```{r}
nrow(mydata)
```

  - Number of events
```{r}
apply(mydata[,c("event_Stroke_total_ind", "event_IHD_total_ind", "event_Stroke_fatal_ind", "event_IHD_fatal_ind", "event_Fatal_CVD_total_ind")], 2, function(x) summary(as.factor(x)))
```

  - Follow-up
```{r}
mean((mydata$censoring_date - mydata$study_date)/365.25)
sd((mydata$censoring_date - mydata$study_date)/365.25)
```

Split FI into quintile groups

```{r fi-process}
mydata$fi_transformed= (mydata$fi.score_full - mean(mydata$fi.score_full))/sd(mydata$fi.score_full) # per 1 SD
mydata$fi_deficit = mydata$fi.score_full/(1/23) # per 1 deficit increment
mydata$fi_0.1 = mydata$fi.score_full/0.1 # per 0.1 unit increment

hist(mydata$fi_transformed)
hist(mydata$fi.score_full)
```

```{r fi-type, eval=FALSE, include=FALSE}
if (fi_type == "original"){
  
  mydata$fi5= factor(ntile(mydata$fi.score_full, 5)) # split FI into quintile groups 
  mydata$ficat = mydata$fi.cat_full

} else if (fi_type == "modified"){
  
  mydata$fi5= factor(ntile(mydata$fi.score_no_cv, 5)) # split FI into quintile groups 
  mydata$ficat = mydata$fi.cat_no_cv
  
} else {
   
}
```

Make outcome data long

```{r make-data-long}
temp = mydata %>% 
  select(csid, event_All_cause_mortality_ind = died, event_All_cause_mortality_date = date_of_death, contains("event_"))

temp = pivot_longer(temp, cols = -1, names_pattern = "(.*)(ind|date)$", names_to = c("names", ".value"))
  # names_pattern = ()()  means look for 2 parts; (.*) means first part should have zero or more characters; (ind|date)$ means the second part should end either as ind or date

temp$names = as.factor(temp$names)
```

Split data on outcome and save as list

```{r split-data}
mylist = split(temp, f = temp$names)
```

## Age-at-risk stratified Cox proportional hazards model

Create function to run analysis

```{r create-function-age-at-risk}
source("J:/R projects/PhD-by-chapter/source codes/ckb resources/expand_age_at_risk.R")

run_age_at_risk_cox_model = function(dataset, frailty, sex){
  
  # Variable input character will not be recognised in dplyr
  x = sym(frailty)
  
  # Merge dob_anon, study_date and rename variables
  dataset = dataset %>% 
    left_join(select(mydata, csid, dob_anon, study_date, !!x, is_female), by = "csid") %>% # mydata contains frailty variable and other covariates
    select(csid, dob_anon, study_date, endpoint = ind, endpoint_date = date, !!x, is_female)
  
  if (sex == "F"){
    
    dataset = dataset[dataset$is_female == 1, ]
    
    formula = as.formula(paste("Surv(time_in, time_out, endpoint) ~ ", frailty," + strata(as.factor(XAgeGrp)) + region_code + household_income + marital_status + occupation + highest_education + hypertension_diag + diabetes_diag + alcohol_category + smoking_group + poor_adiposity + poor_diet + met"))
    
  } else if (sex == "M"){
    
    dataset = dataset[dataset$is_female == 0, ]
    
    formula = as.formula(paste("Surv(time_in, time_out, endpoint) ~ ", frailty," + strata(as.factor(XAgeGrp)) + region_code + household_income + marital_status + occupation + highest_education + hypertension_diag + diabetes_diag + alcohol_category + smoking_group + poor_adiposity + poor_diet + met"))
    
  } else if (sex == "T"){
    
    dataset = dataset
    
    formula = as.formula(paste("Surv(time_in, time_out, endpoint) ~ ", frailty," + strata(as.factor(XAgeGrp)) + is_female + region_code + household_income + marital_status + occupation + highest_education + hypertension_diag + diabetes_diag + alcohol_category + smoking_group + poor_adiposity + poor_diet + met"))
    
  } else {
  }
  
  # Expand data by age-at-risk
  dataset_long = expand_age_at_risk(df = dataset,
                                    ages = c(40, 45, 50, 55, 60, 65, 70, 75, 80))
  #head(dataset_long)
  
  # Merge columns - expanded dataset does not contain the risk factor (frailty) or other covariates
  to_merge = mydata %>% 
    select(csid, !!x, is_female, region_code, household_income, marital_status, occupation, highest_education, met, hypertension_diag, diabetes_diag, alcohol_category, smoking_group, poor_adiposity, poor_diet)
  
  dataset_long = left_join(dataset_long, to_merge, by = "csid")
  
  # Cox proportional hazards model
  fit = survival::coxph(formula, data = dataset_long, ties = "breslow") # the expanded dataframe contains 'XAgeGrp' column which can be used to stratify the CoxPH model, the time columns specify the start and end time for intervals
  #summary(fit)
  
  # Compute onfidence intervals
  hrs = fit %>% 
    tidy(conf.int = TRUE, exponentiate = TRUE) %>% 
    select(term, estimate, starts_with("conf"))

  hrs = hrs[1, ] # save results for frailty

  return(hrs)
  
}  
```

Run function on each outcome

```{r run-age-at-risk-analysis, message=FALSE, error=FALSE, include=FALSE, eval=FALSE}
#mylist1 = mylist[c(3, 9, 24)]
mylist1 = mylist[c("event_Fatal_CVD_total_", "event_IHD_fatal_", "event_Stroke_fatal_", "event_All_cause_mortality_")]

hrlist_fi_trans_t = lapply(mylist1, run_age_at_risk_cox_model, frailty = "fi_transformed", sex = "T")
hrlist_fi_trans_m = lapply(mylist1, run_age_at_risk_cox_model, frailty = "fi_transformed", sex = "M")
hrlist_fi_trans_f = lapply(mylist1, run_age_at_risk_cox_model, frailty = "fi_transformed", sex = "F")

hrlist_fi_deficit_t = lapply(mylist1, run_age_at_risk_cox_model, frailty = "fi_deficit", sex = "T")
hrlist_fi_deficit_m = lapply(mylist1, run_age_at_risk_cox_model, frailty = "fi_deficit", sex = "M")
hrlist_fi_deficit_f = lapply(mylist1, run_age_at_risk_cox_model, frailty = "fi_deficit", sex = "F")

hrlist_fi_0.1_t = lapply(mylist1, run_age_at_risk_cox_model, frailty = "fi_0.1", sex = "T")
hrlist_fi_0.1_m = lapply(mylist1, run_age_at_risk_cox_model, frailty = "fi_0.1", sex = "M")
hrlist_fi_0.1_f = lapply(mylist1, run_age_at_risk_cox_model, frailty = "fi_0.1", sex = "F")
```

```{r}
print(hrlist_fi_deficit_t)
print(hrlist_fi_trans_t)
print(hrlist_fi_0.1_t)
```

Rename and reorder list

```{r rename-and-reorder-list, eval=FALSE, include=FALSE}
# Function to rename and reorder list
rename_reorder_list = function(mylist){
  
  # Rename list using stringr
  names(mylist) = stringr::word(stringr::str_replace_all(names(mylist), "_", " ") # replace all underscore with space
                                  , start=2, end=-2) # extract string from 2 word until last second word
  
  # Reorder list elements
  mylist = mylist[c("Stroke incidence", "Stroke fatal", "Stroke total",
                    "Ischaemic stroke incidence", "Ischaemic stroke fatal", "Ischaemic stroke total",
                    "Haemorrhagic stroke incidence", "Haemorrhagic stroke fatal", "Haemorrhagic stroke total",
                    "IHD incidence", "IHD fatal", "IHD total",
                    "MI incidence", "MI fatal", "MI total",
                    "MVE incidence", "MVE fatal", "MVE total",
                    "MCE incidence", "MCE fatal", "MCE total",
                    "Fatal CVD total", "All cause mortality")]
}

hrlist = list(hrlist_fi5_t, hrlist_fi5_m, hrlist_fi5_f,
              hrlist_ficat_t, hrlist_ficat_m, hrlist_ficat_f)

names(hrlist) = c("HRs FI quintile and outcomes overall", "HRs FI quintile and outcomes in male", "HRs FI quintile and outcomes in female",
                  "HRs FI level and outcomes overall", "HRs FI level and outcomes in male", "HRs FI level and outcomes in female")

hrlist = lapply(hrlist, rename_reorder_list)
```

Save output for plotting later

```{r save-output}
saveRDS(hrlist, file = paste0("J:/R projects/PhD-by-chapter/Chapter6-prospective-association/rds objects/output data/HRs baseline FI (", fi_type, ") and outcomes.rds"))
```