---
title: "Cox regression between baseline frailty and outcomes"
output: html_document
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Required libraries
```{r}
library(ckbplotr)
library(cowplot)
library(Epi)        # for float(), which compute floating absolute risks using the method by Easton et al.
library(kableExtra)
library(knitr)
library(survival)   # for Surv(), cox regression
library(tidyverse)
```

## Open and process data

Participant list for main analysis

```{r}
ptlist_main_analysis = readRDS("J:/R projects/PhD-by-chapter/rds objects/ptlist_main_analysis.rds")
```

Baseline frailty index (excluding cardiovascular deficits)

```{r}
data_baseline_fi_all = readRDS("J:/R projects/ckb-frailty-index/code/analysis/data_baseline_fi_all.rds")

frailty = data_baseline_fi_all %>% 
  select(csid, fi = fi.score_no_cv, ficat = fi.cat_no_cv) %>% 
  filter(csid %in% ptlist_main_analysis)
```

Baseline covariates

```{r}
covariates = readRDS("J:/R projects/PhD-by-chapter/rds objects/covariates_for_main_cohort.rds")
```

Outcome data

```{r}
events = readRDS("J:/R projects/PhD-by-chapter/rds objects/cvd_event_for_main_cohort.rds")
```

Merge data

```{r}
mydata = events %>% 
  left_join(covariates, by = "csid") %>% 
  left_join(frailty, by = "csid")

mydata = mydata %>% 
  mutate_at(c("region_code", "is_female", "household_income", "marital_status", "occupation", "highest_education", "hypertension_diag", "diabetes_diag", "alcohol_category", "smoking_group", "poor_adiposity", "self_rated_health", "poor_diet"), ~as.factor(.))

str(mydata)
```

Subset data

```{r}
if (analysis == "no_death_3y"){
  
  mydata$days_from_entry_to_death = mydata$date_of_death - mydata$study_date # calculate number of days from study entry to death
  
  mydata$death_within_3y_entry = if_else(mydata$days_from_entry_to_death < 365.25*3, 1, 0) # index days less than 3 years
  
  mydata = mydata %>% filter(is.na(death_within_3y_entry) | death_within_3y_entry==0) %>% # select participants alive or died after 3 years
    select(-days_from_entry_to_death, -death_within_3y_entry) # remove variables
  
} else if (analysis == "no_cvd_3y"){
  
  temp = mydata %>% 
    select(csid, study_date, contains("total_date"), contains("incidence_date"), contains("fatal_date")) %>% 
    mutate(
      across( .cols = c(contains("total_date"), contains("incidence_date"), contains("fatal_date")), ~ .x - study_date) # calculate the time between study entry date and date of event
    ) %>% 
    mutate(
      across( .cols = c(contains("total_date"), contains("incidence_date"), contains("fatal_date")), ~ if_else(.x < 365.25*3, 1, 0)) # index events occurring less than 3 years since study entry
    ) 
  
  temp$index = rowSums(temp[, -c(1:2)], na.rm = TRUE) # index participants with any CV events within 3 years of follow-up
  temp = temp[temp$index>=1, "csid"] # select these participant IDs
  
  mydata = mydata %>% anti_join(temp, by = "csid") # anti_join returns all rows from first table which does not match values from the second
  
} else if (analysis == "no_poor_srh"){
  
  mydata = mydata %>% filter(!self_rated_health==3) # remove participants who have poor self-rated health at baseline
  
  mydata$self_rated_health = droplevels(mydata$self_rated_health) # drop unused factor levels
  
} else {
  
}
```

Number of participants included:

  - `r analysis` analysis: `r nrow(mydata)`

Process data

```{r, message=FALSE, warning=FALSE}
mydata = mydata %>% 
  mutate_at(vars(contains("_ind")), ~replace(., is.na(.), 0)) # replace missing values with zeros for dummy variables 

mydata = mydata %>% 
  mutate_at(vars(contains("total_date"), contains("incidence_date"), contains("fatal_date")), ~replace(., is.na(.), censoring_date)) %>% # if empty dates (i.e. no event), then replace empty values with censoring date
  mutate(
    across( .cols = c(contains("total_date"), contains("incidence_date"), contains("fatal_date")), ~ .x - study_date) # calculate the time between study entry date and date of event/censoring
    )

mydata$fi5= factor(ntile(mydata$fi, 5)) # split FI into quintile groups 
  
kable(head(mydata)) %>% kableExtra::kable_styling()
```

```{r, message=FALSE, warning=FALSE}
mydata %>%
  ggplot( aes(x=fi, fill=ficat)) +
    geom_histogram( color="#e9ecef", binwidth=0.01, alpha=0.6, position = 'identity') +
    #scale_fill_gradient() +
    theme_minimal() +
    labs(fill="",
         x = "FI score",
         y = "Count")

mydata %>%
  ggplot( aes(x=fi, fill=fi5)) +
    geom_histogram( color="#e9ecef", binwidth=0.01, alpha=0.6, position = 'identity') +
    #scale_fill_gradient() +
    theme_minimal() +
    labs(fill="",
         x = "FI score",
         y = "Count")
```

## Prospective association between frailty index (FI) and various cardiovascular events

```{r, eval = FALSE, include = FALSE}
# Test code for cox model 
model = survival::coxph(Surv(event_stroke_fatal_date, event_stroke_fatal_ind) ~ ficat + age_at_study_date_x100 + is_female + as.factor(region_code), data = mydata)
temp = Epi::float(model)
tab = table(mydata[["ficat"]], mydata[["event_stroke_fatal_ind"]])

model = survival::coxph(Surv(event_stroke_fatal_date, event_stroke_fatal_ind) ~ fi5 + age_at_study_date_x100 + is_female + as.factor(region_code), data = mydata)
temp = Epi::float(model)
tab = table(mydata[["fi5"]], mydata[["event_stroke_fatal_ind"]])
```

```{r coxph}
# Function to run cox models adjusted for age, sex, and region
func_coxph_floating_var = function(outcome, frailty){
  
  # set up variables as character
  x = frailty
  y_date = paste0("event_", outcome, "_date")
  y_ind = paste0("event_", outcome, "_ind")
  
  # formula
  formula = as.formula(paste("Surv(", y_date, ", ", y_ind, ") ~ ", x," + age_at_study_date_x100 + is_female + region_code"))

  # model
  model = survival::coxph(formula, data = mydata)
  temp = Epi::float(model)
  
  # numbers
  tab = table(mydata[[x]], mydata[[y_ind]])

  # Extract data
  results = data.frame(
    est = temp$coef,
    se = sqrt(temp$var),
    rf = names(temp$coef),
    n = tab[1:dim(tab)[1], 2] 
  )
                       
  return(results)
  
}
```

```{r}
# Function to run cox models adjusted for age, sex, and region
func_coxph_floating_var2 = function(outcome, frailty){
  
  # set up variables as character
  x = frailty
  y_date = paste0("event_", outcome, "_date")
  y_ind = paste0("event_", outcome, "_ind")
  
  # formula
  formula = as.formula(paste("Surv(", y_date, ", ", y_ind, ") ~ ", x," + age_at_study_date_x100 + is_female + region_code + household_income + marital_status + occupation + highest_education + hypertension_diag + diabetes_diag + alcohol_category + smoking_group + poor_adiposity + met"))

  # model
  model = survival::coxph(formula, data = mydata)
  temp = Epi::float(model)
  
  # numbers
  tab = table(mydata[[x]], mydata[[y_ind]])

  # Extract data
  results = data.frame(
    est = temp$coef,
    se = sqrt(temp$var),
    rf = names(temp$coef),
    n = tab[1:dim(tab)[1], 2] 
  )
                       
  return(results)
  
}
```

```{r}
# List outcomes
list_outcomes = list("Stroke_incidence", "Stroke_fatal", "Stroke_total", 
                     "Ischaemic_stroke_incidence", "Ischaemic_stroke_fatal", "Ischaemic_stroke_total", 
                     "Haemorrhagic_stroke_incidence", "Haemorrhagic_stroke_fatal", "Haemorrhagic_stroke_total", 
                     "IHD_incidence", "IHD_fatal", "IHD_total",
                     "MI_incidence", "MI_fatal", "MI_total", 
#TIA_incidence", "TIA_fatal", "TIA_total",
                     "MVE_incidence", "MVE_fatal", "MVE_total",
                     "MCE_incidence", "MCE_fatal", "MCE_total",
                     "Fatal_CVD_total")

# Run cox models for outcomes by frailty categories
list_input = rep(list("ficat"), length(list_outcomes)) # number of events studied

# Minimally adjuste
cox_results1 = purrr::map2(list_outcomes, list_input, func_coxph_floating_var)

names(cox_results1) = paste(list_outcomes, list_input)

# Fully adjusted for confounders
cox_results1a = purrr::map2(list_outcomes, list_input, func_coxph_floating_var2)

names(cox_results1a) = paste(list_outcomes, list_input)

# Run cox models for outcomes by frailty categories
list_input = rep(list("fi5"), length(list_outcomes)) # number of events studied

# Minimally adjusted
cox_results2 = purrr::map2(list_outcomes, list_input, func_coxph_floating_var)

names(cox_results2) = paste(list_outcomes, list_input)

# Fully adjusted for confounders
cox_results2a = purrr::map2(list_outcomes, list_input, func_coxph_floating_var2)

names(cox_results2a) = paste(list_outcomes, list_input)
```

### Cox regression of outcomes by frailty categories

```{r}
# Function to make first letters of a word upper case
firstup = function(x) {
  substr(x, 1, 1) = toupper(substr(x, 1, 1))
  x
}
```

```{r, eval=FALSE, include=FALSE}
# testing

paneled_shape_plot = function(x) {
  plots = list()
  
  for (i in 1:length(x)){
    
    title = stringr::word(names(x[i]), 1) %>% # extract first word (i.e. all string up to first space)
    str_replace("_", " ") %>% # replace underscores with a space
    firstup() # make first letter uppercase
    
    x[[i]]$rf = factor(x[[i]]$rf, levels = c("Non-frail", "Pre-frail", "Frail"))
    
    plots[[i]] = ckbplotr::make_shape_plot(x[[i]],
                            col.x          = "rf",
                            col.estimate   = "est",
                            col.stderr     = "se",
                            col.n          = "n",
                            exponentiate   = TRUE, 
                            xlims          = c(1, 3),
                            ylims          = c(0.85, 2.5),
                            xlab           = "",
                            ylab           = "Hazard ratio (95% CI)",
                            scalepoints    = TRUE,
                            title          = title,
                            printplot      = FALSE, 
                            showcode = FALSE)
    
  }
}

paneled_shape_plot(cox_results1)
```


```{r, out.width = "100%"}
plots1 = list()

for (i in 1:length(cox_results1)){
  
  title = stringr::word(names(cox_results1[i]), 1) %>% # extract first word (i.e. all string up to first space)
    str_replace_all("_", " ") %>% # replace underscores with a space
    firstup() # make first letter uppercase  
  
  cox_results1[[i]]$rf = factor(cox_results1[[i]]$rf, levels = c("Non-frail", "Pre-frail", "Frail"))
  
  plots1[[i]] = ckbplotr::make_shape_plot(cox_results1[[i]],
                            col.x          = "rf",
                            col.estimate   = "est",
                            col.stderr     = "se",
                            col.n          = "n",
                            exponentiate   = TRUE, 
                            xlims          = c(1, 3),
                            ylims          = c(0.85, 2.5),
                            xlab           = "",
                            ylab           = "Hazard ratio (95% CI)",
                            scalepoints    = TRUE,
                            title          = title,
                            printplot      = FALSE, 
                            showcode = FALSE)
  
}

plot1a = cowplot::plot_grid(plots1[[1]]$plot, plots1[[2]]$plot, plots1[[3]]$plot, 
                           plots1[[4]]$plot, plots1[[5]]$plot, plots1[[6]]$plot,
                           plots1[[7]]$plot, plots1[[8]]$plot, plots1[[9]]$plot,
                           ncol=3, nrow=3, align = "hv")

plot1b = cowplot::plot_grid(plots1[[10]]$plot, plots1[[11]]$plot, plots1[[12]]$plot, 
                           plots1[[13]]$plot, plots1[[14]]$plot, plots1[[15]]$plot,
                           ncol=3, nrow=3, align = "hv")

plot1c = cowplot::plot_grid(plots1[[16]]$plot, plots1[[17]]$plot, plots1[[18]]$plot, 
                           plots1[[19]]$plot, plots1[[20]]$plot, plots1[[21]]$plot,
                           plots1[[22]]$plot,
                           ncol=3, nrow=3, align = "hv")

plot1a
plot1b
plot1c

cowplot::save_plot(paste0("J:/R projects/PhD-by-chapter/Chapter6-prospective-association/outputs/HRplots_CVD1_by_frailty_level_", analysis, ".png"), plot1a, base_height = 8, base_aspect_ratio = 1)
cowplot::save_plot(paste0("J:/R projects/PhD-by-chapter/Chapter6-prospective-association/outputs/HRplots_CVD2_by_frailty_level_", analysis, ".png"), plot1b, base_height = 8, base_aspect_ratio = 1)
cowplot::save_plot(paste0("J:/R projects/PhD-by-chapter/Chapter6-prospective-association/outputs/HRplots_CVD3_by_frailty_level_", analysis, ".png"), plot1c, base_height = 8, base_aspect_ratio = 1)
```
Cox proportional hazards models adjusted for age, sex and region.

### Cox regression of outcomes by frailty quintile groups

```{r, out.width = "100%"}
plots2 = list()

for (i in 1:length(cox_results2)){
  
  title = stringr::word(names(cox_results2[i]), 1) %>% # extract first word (i.e. all string up to first space)
    str_replace_all("_", " ") %>% # replace underscores with a space
    firstup() # make first letter uppercase
  
  plots2[[i]] = ckbplotr::make_shape_plot(cox_results2[[i]],
                            col.x          = "rf",
                            col.estimate   = "est",
                            col.stderr     = "se",
                            col.n          = "n",
                            exponentiate   = TRUE, 
                            xlims          = c(1, 5),
                            ylims          = c(0.85, 2.2),
                            xlab           = "",
                            ylab           = "Hazard ratio (95% CI)",
                            scalepoints    = TRUE,
                            title          = title,
                            printplot      = FALSE, 
                            showcode = FALSE)
}  

plot2a = cowplot::plot_grid(plots2[[1]]$plot, plots2[[2]]$plot, plots2[[3]]$plot, 
                           plots2[[4]]$plot, plots2[[5]]$plot, plots2[[6]]$plot,
                           plots2[[7]]$plot, plots2[[8]]$plot, plots2[[9]]$plot,
                           ncol=3, nrow=3, align = "hv")

plot2b = cowplot::plot_grid(plots2[[10]]$plot, plots2[[11]]$plot, plots2[[12]]$plot, 
                           plots2[[13]]$plot, plots2[[14]]$plot, plots2[[15]]$plot,
                           ncol=3, nrow=3, align = "hv")

plot2c = cowplot::plot_grid(plots2[[16]]$plot, plots2[[17]]$plot, plots2[[18]]$plot, 
                           plots2[[19]]$plot, plots2[[20]]$plot, plots2[[21]]$plot,
                           plots2[[22]]$plot,
                           ncol=3, nrow=3, align = "hv")

plot2a
plot2b
plot2c


cowplot::save_plot(paste0("J:/R projects/PhD-by-chapter/Chapter6-prospective-association/outputs/HRplots_CVD1_by_frailty_quintile_", analysis, ".png"), plot2a, base_height = 8, base_aspect_ratio = 1)
cowplot::save_plot(paste0("J:/R projects/PhD-by-chapter/Chapter6-prospective-association/outputs/HRplots_CVD2_by_frailty_quintile_", analysis, ".png"), plot2b, base_height = 8, base_aspect_ratio = 1)
cowplot::save_plot(paste0("J:/R projects/PhD-by-chapter/Chapter6-prospective-association/outputs/HRplots_CVD3_by_frailty_quintile_", analysis, ".png"), plot2c, base_height = 8, base_aspect_ratio = 1)
```

Cox proportional hazards models adjusted for age, sex and region.

#### The same plots with full adjustment of confounders

```{r, out.width = "100%"}
plots1 = list()

for (i in 1:length(cox_results1a)){
  
  title = stringr::word(names(cox_results1a[i]), 1) %>% # extract first word (i.e. all string up to first space)
    str_replace("_", " ") %>% # replace underscores with a space
    firstup() # make first letter uppercase  
  
  cox_results1a[[i]]$rf = factor(cox_results1a[[i]]$rf, levels = c("Non-frail", "Pre-frail", "Frail"))
  
  plots1[[i]] = ckbplotr::make_shape_plot(cox_results1a[[i]],
                            col.x          = "rf",
                            col.estimate   = "est",
                            col.stderr     = "se",
                            col.n          = "n",
                            exponentiate   = TRUE, 
                            xlims          = c(1, 3),
                            ylims          = c(0.85, 2.5),
                            xlab           = "",
                            ylab           = "Hazard ratio (95% CI)",
                            scalepoints    = TRUE,
                            title          = title,
                            printplot      = FALSE, 
                            showcode = FALSE)
  
}

plot1a = cowplot::plot_grid(plots1[[1]]$plot, plots1[[2]]$plot, plots1[[3]]$plot, 
                           plots1[[4]]$plot, plots1[[5]]$plot, plots1[[6]]$plot,
                           plots1[[7]]$plot, plots1[[8]]$plot, plots1[[9]]$plot,
                           ncol=3, nrow=3, align = "hv")

plot1b = cowplot::plot_grid(plots1[[10]]$plot, plots1[[11]]$plot, plots1[[12]]$plot, 
                           plots1[[13]]$plot, plots1[[14]]$plot, plots1[[15]]$plot,
                           ncol=3, nrow=3, align = "hv")

plot1c = cowplot::plot_grid(plots1[[16]]$plot, plots1[[17]]$plot, plots1[[18]]$plot, 
                           plots1[[19]]$plot, plots1[[20]]$plot, plots1[[21]]$plot,
                           plots1[[22]]$plot,
                           ncol=3, nrow=3, align = "hv")

plot1a
plot1b
plot1c

cowplot::save_plot(paste0("J:/R projects/PhD-by-chapter/Chapter6-prospective-association/outputs/HRplots_CVD1_by_frailty_level_fulladj_", analysis, ".png"), plot1a, base_height = 8, base_aspect_ratio = 1)
cowplot::save_plot(paste0("J:/R projects/PhD-by-chapter/Chapter6-prospective-association/outputs/HRplots_CVD2_by_frailty_level_fulladj_", analysis, ".png"), plot1b, base_height = 8, base_aspect_ratio = 1)
cowplot::save_plot(paste0("J:/R projects/PhD-by-chapter/Chapter6-prospective-association/outputs/HRplots_CVD3_by_frailty_level_fulladj_", analysis, ".png"), plot1c, base_height = 8, base_aspect_ratio = 1)
```
Cox proportional hazards models adjusted for age, sex, region, income, marital status, occupation, education level, hypertension diagnosis, diabetes diagnosis, alcohol category, smoking category, poor adiposity and physical activity. 

```{r, out.width = "100%"}
plots2 = list()

for (i in 1:length(cox_results2a)){
  
  title = stringr::word(names(cox_results2a[i]), 1) %>% # extract first word (i.e. all string up to first space)
    str_replace("_", " ") %>% # replace underscores with a space
    firstup() # make first letter uppercase
  
  plots2[[i]] = ckbplotr::make_shape_plot(cox_results2a[[i]],
                            col.x          = "rf",
                            col.estimate   = "est",
                            col.stderr     = "se",
                            col.n          = "n",
                            exponentiate   = TRUE, 
                            xlims          = c(1, 5),
                            ylims          = c(0.85, 2.2),
                            xlab           = "",
                            ylab           = "Hazard ratio (95% CI)",
                            scalepoints    = TRUE,
                            title          = title,
                            printplot      = FALSE, 
                            showcode = FALSE)
}  

plot2a = cowplot::plot_grid(plots2[[1]]$plot, plots2[[2]]$plot, plots2[[3]]$plot, 
                           plots2[[4]]$plot, plots2[[5]]$plot, plots2[[6]]$plot,
                           plots2[[7]]$plot, plots2[[8]]$plot, plots2[[9]]$plot,
                           ncol=3, nrow=3, align = "hv")

plot2b = cowplot::plot_grid(plots2[[10]]$plot, plots2[[11]]$plot, plots2[[12]]$plot, 
                           plots2[[13]]$plot, plots2[[14]]$plot, plots2[[15]]$plot,
                           ncol=3, nrow=3, align = "hv")

plot2c = cowplot::plot_grid(plots2[[16]]$plot, plots2[[17]]$plot, plots2[[18]]$plot, 
                           plots2[[19]]$plot, plots2[[20]]$plot, plots2[[21]]$plot,
                           plots2[[22]]$plot,
                           ncol=3, nrow=3, align = "hv")

plot2a
plot2b
plot2c


cowplot::save_plot(paste0("J:/R projects/PhD-by-chapter/Chapter6-prospective-association/outputs/HRplots_CVD1_by_frailty_quintile_fulladj_", analysis, ".png"), plot2a, base_height = 8, base_aspect_ratio = 1)
cowplot::save_plot(paste0("J:/R projects/PhD-by-chapter/Chapter6-prospective-association/outputs/HRplots_CVD2_by_frailty_quintile_fulladj_", analysis, ".png"), plot2b, base_height = 8, base_aspect_ratio = 1)
cowplot::save_plot(paste0("J:/R projects/PhD-by-chapter/Chapter6-prospective-association/outputs/HRplots_CVD3_by_frailty_quintile_fulladj_", analysis, ".png"), plot2c, base_height = 8, base_aspect_ratio = 1)
```
Cox proportional hazards models adjusted for age, sex, region, income, marital status, occupation, education level, hypertension diagnosis, diabetes diagnosis, alcohol category, smoking category, poor adiposity and physical activity. 