---
title: "Sensitivity analysis results"
output: html_document
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Libraries

```{r}
library(data.table)
```

## Sensitivity analysis 

### Association between frailty and outcomes by length of follow-up

Open saved data

```{r}
hrlist = readRDS(file = "J:/R projects/PhD-by-chapter/Chapter6-prospective-association/rds objects/output data/HRs baseline FI and outcomes (sensitivity years of follow-up).rds")
```

Each element of list contains:

```{r}
names(hrlist)
```
Each element contains HRs for frailty an the following outcomes:

```{r}
names(hrlist[[1]])
```
Each result data looks like:

```{r}
hrlist[[1]][[1]]
```

Process data

```{r}
mfi1_stroke = data.table::rbindlist(list("1" = hrlist[[1]][[1]], 
                                         "3" = hrlist[[2]][[1]],
                                         "5" = hrlist[[3]][[1]], 
                                         "8" = hrlist[[4]][[1]]), idcol = "followup")
mfi1_ihd = data.table::rbindlist(list("1" = hrlist[[1]][[2]], 
                                         "3" = hrlist[[2]][[2]],
                                         "5" = hrlist[[3]][[2]], 
                                         "8" = hrlist[[4]][[2]]), idcol = "followup")
mfi1_cvd = data.table::rbindlist(list("1" = hrlist[[1]][[3]], 
                                         "3" = hrlist[[2]][[3]],
                                         "5" = hrlist[[3]][[3]], 
                                         "8" = hrlist[[4]][[3]]), idcol = "followup")
mfi1_death = data.table::rbindlist(list("1" = hrlist[[1]][[4]], 
                                         "3" = hrlist[[2]][[4]],
                                         "5" = hrlist[[3]][[4]], 
                                         "8" = hrlist[[4]][[4]]), idcol = "followup")
mfi3_stroke = data.table::rbindlist(list("1" = hrlist[[5]][[1]], 
                                         "3" = hrlist[[6]][[1]],
                                         "5" = hrlist[[7]][[1]], 
                                         "8" = hrlist[[8]][[1]]), idcol = "followup")
mfi3_ihd = data.table::rbindlist(list("1" = hrlist[[5]][[2]], 
                                         "3" = hrlist[[6]][[2]],
                                         "5" = hrlist[[7]][[2]], 
                                         "8" = hrlist[[8]][[2]]), idcol = "followup")
mfi3_cvd = data.table::rbindlist(list("1" = hrlist[[5]][[3]], 
                                         "3" = hrlist[[6]][[3]],
                                         "5" = hrlist[[7]][[3]], 
                                         "8" = hrlist[[8]][[3]]), idcol = "followup")
mfi3_death = data.table::rbindlist(list("1" = hrlist[[5]][[4]], 
                                         "3" = hrlist[[6]][[4]],
                                         "5" = hrlist[[7]][[4]], 
                                         "8" = hrlist[[8]][[4]]), idcol = "followup")
process_data = function(df){
  df = df %>% 
   mutate(hr = format(round(estimate, 2), nsmall=2),
          lq = format(round(conf.low, 2), nsmall=2),
          uq = format(round(conf.high, 2), nsmall=2)) %>% 
   mutate(hrs = paste0(hr, " (", lq, "-", uq, ")")) %>% 
   select(followup, n, hrs)

  return(df)
}

mylist_data = list(mfi1_stroke, mfi1_ihd, mfi1_cvd, mfi1_death,
                   mfi3_stroke, mfi3_ihd, mfi3_cvd, mfi3_death)

processed_data = lapply(mylist_data, process_data)
```

Combine and save data

```{r}
names(processed_data[[1]]) = c("followup", "n.stroke", "hrs.stroke")
names(processed_data[[5]]) = c("followup", "n.stroke", "hrs.stroke")
names(processed_data[[2]]) = c("followup", "n.ihd", "hrs.ihd")
names(processed_data[[6]]) = c("followup", "n.ihd", "hrs.ihd")
names(processed_data[[3]]) = c("followup", "n.cvd", "hrs.cvd")
names(processed_data[[7]]) = c("followup", "n.cvd", "hrs.cvd")
names(processed_data[[4]]) = c("followup", "n.death", "hrs.death")
names(processed_data[[8]]) = c("followup", "n.death", "hrs.death")

temp1 = purrr::reduce(list(processed_data[[1]], processed_data[[2]], processed_data[[3]], processed_data[[4]]), dplyr::left_join, by = "followup")
temp2 = purrr::reduce(list(processed_data[[5]], processed_data[[6]], processed_data[[7]], processed_data[[8]]), dplyr::left_join, by = "followup")

output = bind_rows(list("mFI-1" = temp1, "mFI-3" = temp2), .id = "fi")

output
```

Save output

```{r}
write.csv(output, file = "J:/R projects/PhD-by-chapter/Chapter6-prospective-association/outputs/sensitivity/associations-frailty-outcomes-by-followup-length.csv")
```

### Association between frailty and outcomes by exclusions

Open saved data

```{r}
hrlist = readRDS(file = "J:/R projects/PhD-by-chapter/Chapter6-prospective-association/rds objects/output data/HRs baseline FI and outcomes (sensitivity exclusions).rds")
```

Each element of list contains:

```{r}
names(hrlist)
```
Each element contains HRs for frailty an the following outcomes:

```{r}
names(hrlist[[1]])
```
Each result data looks like:

```{r}
hrlist[[1]][[1]]
```

Process data

```{r}
mfi1_stroke = data.table::rbindlist(list("neversmokers" = hrlist[[4]][[1]], 
                                         "noweightloss" = hrlist[[2]][[1]],
                                         "nopoorselfra" = hrlist[[3]][[1]], 
                                         "noearlydeath" = hrlist[[1]][[1]]), idcol = "exclusion")
mfi1_ihd =    data.table::rbindlist(list("neversmokers" = hrlist[[4]][[2]], 
                                         "noweightloss" = hrlist[[2]][[2]],
                                         "nopoorselfra" = hrlist[[3]][[2]], 
                                         "noearlydeath" = hrlist[[1]][[2]]), idcol = "exclusion")
mfi1_cvd =    data.table::rbindlist(list("neversmokers" = hrlist[[4]][[3]], 
                                         "noweightloss" = hrlist[[2]][[3]],
                                         "nopoorselfra" = hrlist[[3]][[3]], 
                                         "noearlydeath" = hrlist[[1]][[3]]), idcol = "exclusion")
mfi1_death =  data.table::rbindlist(list("neversmokers" = hrlist[[4]][[4]], 
                                         "noweightloss" = hrlist[[2]][[4]],
                                         "nopoorselfra" = hrlist[[3]][[4]], 
                                         "noearlydeath" = hrlist[[1]][[4]]), idcol = "exclusion")
mfi3_stroke = data.table::rbindlist(list("neversmokers" = hrlist[[8]][[1]], 
                                         "noweightloss" = hrlist[[6]][[1]],
                                         "nopoorselfra" = hrlist[[7]][[1]], 
                                         "noearlydeath" = hrlist[[5]][[1]]), idcol = "exclusion")
mfi3_ihd =    data.table::rbindlist(list("neversmokers" = hrlist[[8]][[2]], 
                                         "noweightloss" = hrlist[[6]][[2]],
                                         "nopoorselfra" = hrlist[[7]][[2]], 
                                         "noearlydeath" = hrlist[[5]][[2]]), idcol = "exclusion")
mfi3_cvd =    data.table::rbindlist(list("neversmokers" = hrlist[[8]][[3]], 
                                         "noweightloss" = hrlist[[6]][[3]],
                                         "nopoorselfra" = hrlist[[7]][[3]], 
                                         "noearlydeath" = hrlist[[5]][[3]]), idcol = "exclusion")
mfi3_death =  data.table::rbindlist(list("neversmokers" = hrlist[[8]][[4]], 
                                         "noweightloss" = hrlist[[6]][[4]],
                                         "nopoorselfra" = hrlist[[7]][[4]], 
                                         "noearlydeath" = hrlist[[5]][[4]]), idcol = "exclusion")

process_data = function(df){
  df = df %>% 
   mutate(hr = format(round(estimate, 2), nsmall=2),
          lq = format(round(conf.low, 2), nsmall=2),
          uq = format(round(conf.high, 2), nsmall=2)) %>% 
   mutate(hrs = paste0(hr, " (", lq, "-", uq, ")")) %>% 
   select(exclusion, n, hrs)

  return(df)
}

mylist_data = list(mfi1_stroke, mfi1_ihd, mfi1_cvd, mfi1_death,
                   mfi3_stroke, mfi3_ihd, mfi3_cvd, mfi3_death)

processed_data = lapply(mylist_data, process_data)
```

Combine and save data

```{r}
names(processed_data[[1]]) = c("exclusion", "n.stroke", "hrs.stroke")
names(processed_data[[5]]) = c("exclusion", "n.stroke", "hrs.stroke")
names(processed_data[[2]]) = c("exclusion", "n.ihd", "hrs.ihd")
names(processed_data[[6]]) = c("exclusion", "n.ihd", "hrs.ihd")
names(processed_data[[3]]) = c("exclusion", "n.cvd", "hrs.cvd")
names(processed_data[[7]]) = c("exclusion", "n.cvd", "hrs.cvd")
names(processed_data[[4]]) = c("exclusion", "n.death", "hrs.death")
names(processed_data[[8]]) = c("exclusion", "n.death", "hrs.death")

temp1 = purrr::reduce(list(processed_data[[1]], processed_data[[2]], processed_data[[3]], processed_data[[4]]), dplyr::left_join, by = "exclusion")
temp2 = purrr::reduce(list(processed_data[[5]], processed_data[[6]], processed_data[[7]], processed_data[[8]]), dplyr::left_join, by = "exclusion")

output = bind_rows(list("mFI-1" = temp1, "mFI-3" = temp2), .id = "fi")

output
```

Save output

```{r}
write.csv(output, file = "J:/R projects/PhD-by-chapter/Chapter6-prospective-association/outputs/sensitivity/associations-frailty-outcomes-by-exclusions.csv")
```
