---
title: "Sensitivity analysis: reverse causation bias by length of follow-up"
output: html_document
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Sensitivity analysis to explore reverse causation bias by length of follow-up

In the associations between frailty and incident CVD, associations may be due to underlying illness or advanced stage of disease in those with CVD. To explore this, sensitivity analysis will be carried out comparing the strength of associations by length of follow-up.

 - Outcomes within 1 year of study entry
 - Outcomes within 2 years of study entry
 - Outcomes within 4 years of study entry 
 - Outcomes within 7 years of study entry 
 
If the strength of associations attenuate with length of follow-up, this suggests presence of reverse causation bias.
 
#### Required libraries

```{r libraries}
library(broom)
# required for age-at-risk stratified CoxPH model
library(Epi)
library(lubridate)
library(survival)   
library(tidyverse)
```

#### Open and process data

```{r open-data}
# Source script that creates working dataset that includes baseline frailty scores and CVD outcomes with key baseline covariates
source("J:/R projects/PhD-by-chapter/Chapter6-prospective-association/code/source/script to open and process dataset to analyse baseline frailty and outcomes.R")

mydata = mydata %>% 
  mutate(across(contains("_ind"), ~replace(., is.na(.), 0))) # replace missing values with zeros for dummy variables 

mydata = mydata %>% 
  mutate(across(c(contains("total_date"), contains("incidence_date"), contains("fatal_date"), date_of_death), ~replace(., is.na(.), censoring_date))) # if empty dates (i.e. no event), then replace empty values with censoring date
```

Process data

```{r process-data}
process_data = function(df){
  
  temp = df %>% 
    select(csid, study_date, event_All_cause_mortality_ind = died, event_All_cause_mortality_date = date_of_death, event_Stroke_total_ind, event_Stroke_total_date, event_IHD_total_ind, event_IHD_total_date, event_Fatal_CVD_total_ind, event_Fatal_CVD_total_date)
  
  temp = pivot_longer(temp, cols = -c(1:2), names_pattern = "(.*)(ind|date)$", names_to = c("names", ".value"))
    # names_pattern = ()()  means look for 2 parts; (.*) means first part should have zero or more characters; (ind|date)$ means the second part should end either as ind or date
  
  temp$names = as.factor(temp$names)
  
  mylist = split(temp, f = temp$names)
                 
  return(mylist)
}

mylist = process_data(mydata)
```

Create separate datasets by length of follow-up

```{r}
change_followup_length = function(df, n){ # n = length of follow-up to limit the df to
  
  df = df %>% 
    select(csid, study_date, names, old_ind = ind, old_date = date) %>% 
    mutate(temp_date = study_date %m+% years(n)) %>% 
    mutate(ind = if_else(old_date<=temp_date, old_ind, 0),
           date = if_else(old_date<=temp_date, old_date, temp_date)) %>% 
    select(csid, names, date, ind)
  
  return(df)
}

mylist_1y = lapply(mylist, change_followup_length, n=1)
mylist_3y = lapply(mylist, change_followup_length, n=3)
mylist_5y = lapply(mylist, change_followup_length, n=5)
mylist_8y = lapply(mylist, change_followup_length, n=8)
```

### Run age-at-risk Cox PH models

```{r create-function-age-at-risk}
source("J:/R projects/PhD-by-chapter/source codes/ckb resources/expand_age_at_risk.R")

run_age_at_risk_cox_model = function(dataset, fitype){
  
  # Save number of outcomes
  n_outcome = table(dataset$ind)[["1"]]
  
  # Merge dob_anon, study_date and rename variables
  dataset = dataset %>% 
    left_join(select(mydata, csid, dob_anon, study_date, is_female), by = "csid") %>% # mydata contains frailty variable and other covariates
    select(csid, dob_anon, study_date, endpoint = ind, endpoint_date = date, is_female)
  
  formula = as.formula(paste("Surv(time_in, time_out, endpoint) ~ fi_transformed + strata(as.factor(XAgeGrp)) + is_female + region_code + household_income + marital_status + occupation + highest_education + hypertension_diag + diabetes_diag + alcohol_category + smoking_group + fat_body_mass_kg + poor_diet + met"))
  
  # Expand data by age-at-risk
  dataset_long = expand_age_at_risk(df = dataset,
                                    ages = c(40, 45, 50, 55, 60, 65, 70, 75, 80))
  #head(dataset_long)
  
  # Merge columns - expanded dataset does not contain the risk factor (frailty) or other covariates
  to_merge = mydata %>% 
    select(csid, {{fitype}}, is_female, region_code, household_income, marital_status, occupation, highest_education, met, hypertension_diag, diabetes_diag, alcohol_category, smoking_group, fat_body_mass_kg, poor_diet) %>% 
    mutate(fi_transformed = scale({{fitype}})) # transform FI
  
  dataset_long = left_join(dataset_long, to_merge, by = "csid")
  
  # Cox proportional hazards model
  fit = survival::coxph(formula, data = dataset_long, ties = "breslow") # the expanded dataframe contains 'XAgeGrp' column which can be used to stratify the CoxPH model, the time columns specify the start and end time for intervals
  #summary(fit)

  # Compute onfidence intervals
  results = fit %>% 
    tidy(conf.int = TRUE, exponentiate = TRUE) %>% 
    select(term, estimate, starts_with("conf"))
  
  results = results[1, ] # save results for frailty

  results$n = n_outcome # add number of outcomes 
  
  return(results)
  
}  
```


Run function on each outcome

```{r run-age-at-risk-analysis, message=FALSE, error=FALSE}
hrlist_mfi1sd_sty_1 = lapply(mylist_1y, run_age_at_risk_cox_model, fitype = fi.score_no_cv)
hrlist_mfi1sd_sty_3 = lapply(mylist_3y, run_age_at_risk_cox_model, fitype = fi.score_no_cv)
hrlist_mfi1sd_sty_5 = lapply(mylist_5y, run_age_at_risk_cox_model, fitype = fi.score_no_cv)
hrlist_mfi1sd_sty_8 = lapply(mylist_8y, run_age_at_risk_cox_model, fitype = fi.score_no_cv)

hrlist_mfi3sd_sty_1 = lapply(mylist_1y, run_age_at_risk_cox_model, fitype = fi.score_no_21)
hrlist_mfi3sd_sty_3 = lapply(mylist_3y, run_age_at_risk_cox_model, fitype = fi.score_no_21)
hrlist_mfi3sd_sty_5 = lapply(mylist_5y, run_age_at_risk_cox_model, fitype = fi.score_no_21)
hrlist_mfi3sd_sty_8 = lapply(mylist_8y, run_age_at_risk_cox_model, fitype = fi.score_no_21)
```

Rename and reorder list

```{r rename-and-reorder-list}
# Function to rename and reorder list
rename_reorder_list = function(mylist){
  
  # Rename list using stringr
  names(mylist) = stringr::word(stringr::str_replace_all(names(mylist), "_", " ") # replace all underscore with space
                                  , start=2, end=-2) # extract string from 2 word until last second word
  
  # Reorder list elements
  mylist = mylist[c("Stroke total", "IHD total", "Fatal CVD total", "All cause mortality")]
}

hrlist = list(hrlist_mfi1sd_sty_1, hrlist_mfi1sd_sty_3, hrlist_mfi1sd_sty_5, hrlist_mfi1sd_sty_8,
              hrlist_mfi3sd_sty_1, hrlist_mfi3sd_sty_3, hrlist_mfi3sd_sty_5, hrlist_mfi3sd_sty_8)

names(hrlist) = c("HRs mFI-1 1SD and outcomes (1y fu)", 
                  "HRs mFI-1 1SD and outcomes (3y fu)", 
                  "HRs mFI-1 1SD and outcomes (5y fu)", 
                  "HRs mFI-1 1SD and outcomes (8y fu)", 
                  "HRs mFI-3 1SD and outcomes (1y fu)", 
                  "HRs mFI-3 1SD and outcomes (3y fu)", 
                  "HRs mFI-3 1SD and outcomes (5y fu)", 
                  "HRs mFI-3 1SD and outcomes (8y fu)")

hrlist = lapply(hrlist, rename_reorder_list)
```

Save output for plotting later

```{r save-output}
saveRDS(hrlist, file = paste0("J:/R projects/PhD-by-chapter/Chapter6-prospective-association/rds objects/output data/HRs baseline FI and outcomes (sensitivity years of follow-up).rds"))
```
