---
title: "CoxPH subgroup models"
output: html_document
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Required libraries

```{r library}
library(broom)      # for tidy() to get clean HRs and confidence intervals from model fit
library(cowplot)
library(data.table)
library(Epi)        # for float(), which compute floating absolute risks using the method by Easton et al.
library(gridExtra)
library(kableExtra)
library(knitr)
library(lubridate)
library(survival)   # for Surv(), cox regression
library(tidyverse)
```

## Open and process data

Run external script to open and process data

```{r open-data}
# Source script that creates working dataset that includes baseline frailty scores and CVD outcomes with key baseline covariates
source("J:/R projects/PhD-by-chapter/Chapter6-prospective-association/code/source/script to open and process dataset to analyse baseline frailty and outcomes.R")

# open pretty covariates for later
covariates = readRDS("J:/R projects/PhD-by-chapter/rds objects/covariates_pretty_levels_for_everyone.rds")
```

Data structure

```{r data-structure}
str(mydata)
```


Number of participants included:

  - FI analysis: `r nrow(mydata)`

Further process data for age-at-risk analysis to have following variables:

  - csid
  - dob_anon
  - study_date
  - endpoint - binary variable indicating the endpoint (1/0)
  - endpoint_date - date of endpoint event or censoring
  - frailty quintile group
  - outcome_type - indicate outcome type, e.g. stroke, MI, etc.

```{r data-process, message=FALSE, warning=FALSE}
mydata = mydata %>% 
  mutate_at(vars(contains("_ind")), ~replace(., is.na(.), 0)) # replace missing values with zeros for dummy variables 

mydata = mydata %>% 
  mutate_at(vars(contains("total_date"), contains("incidence_date"), contains("fatal_date"), date_of_death), ~replace(., is.na(.), censoring_date)) # if empty dates (i.e. no event), then replace empty values with censoring date

kable(head(mydata)) %>% kableExtra::kable_styling()
```

Transform FI variable

```{r fi-process}
mydata$fi_transformed= (mydata$fi.score_no_cv - mean(mydata$fi.score_no_cv))/sd(mydata$fi.score_no_cv) # per 1 SD
#mydata$fi_deficit = mydata$fi.score_no_cv/(1/23) # per 1 deficit increment
#mydata$fi_0.1 = mydata$fi.score_no_cv/0.1 # per 0.1 unit increment

#hist(mydata$fi_transformed)
#hist(mydata$fi.score_no_cv)
```

Make outcome data long

```{r make-data-long}
temp = mydata %>% 
  select(csid, event_All_cause_mortality_ind = died, event_All_cause_mortality_date = date_of_death, contains("event_")) # rename death variables

temp = pivot_longer(temp, cols = -1, names_pattern = "(.*)(ind|date)$", names_to = c("names", ".value"))
  # names_pattern = ()()  means look for 2 parts; (.*) means first part should have zero or more characters; (ind|date)$ means the second part should end either as ind or date

temp$names = as.factor(temp$names)
```

Split data on outcome and save as list

```{r split-data}
mylist = split(temp, f = temp$names)
```

## Analysis

Select stroke/IHD dataset

```{r}
dataset = mylist[[c("event_Stroke_total_")]]
dataset = mylist[[c("event_IHD_total_")]]
```

Expand data for age-at-risk analysis

```{r}
# Expand data for age-at-risk analysis
source("J:/R projects/PhD-by-chapter/source codes/ckb resources/expand_age_at_risk.R")

# Merge dob_anon, study_date and rename variables
dataset = dataset %>% 
  left_join(select(mydata, csid, dob_anon, study_date), by = "csid") %>% # mydata contains frailty variable and other covariates
  select(csid, dob_anon, study_date, endpoint = ind, endpoint_date = date)

dataset_long = expand_age_at_risk(df = dataset,
                                  ages = c(40, 45, 50, 55, 60, 65, 70, 75, 80))
#head(dataset_long)

# Merge columns - expanded dataset does not contain the risk factor (frailty) or other covariates
to_merge = covariates %>% 
  select(csid, is_female, age_group, region_is_urban, region_code, household_income, marital_status_group, occupation, highest_education, met, hypertension_diag, diabetes_diag, random_glucose_group, sbp_group, fat_body_mass_kg, alcohol_category, smoking_group, poor_diet, sbp_mean, random_glucose) %>% 
  mutate(highest_education = factor(case_when(highest_education == "No formal school" ~ "No formal school",
                                       highest_education == "Primary School" ~ "Primary school",
                                       TRUE ~ "Middle school or higher"), levels = c("No formal school", "Primary school", "Middle school or higher")),
         fat_body_mass_cat = factor(ntile(fat_body_mass_kg, 5), levels=c(1:5)))# recode variables
  
dataset_long = left_join(dataset_long, to_merge, by = "csid") %>% 
  left_join(select(mydata, csid, fi_transformed), by = "csid")
```

Run age-at-risk, sex, and region stratified, where appropriate, Cox proportional hazards models by subgroup

```{r}
# Sex
models_sex = by(dataset_long, dataset_long$is_female, function(x) survival::coxph(Surv(time_in, time_out, endpoint) ~ fi_transformed + strata(as.factor(XAgeGrp)) + strata(region_code) + household_income + marital_status_group + occupation + highest_education + sbp_mean + random_glucose + alcohol_category + smoking_group + poor_diet + fat_body_mass_kg, data = x, ties = "breslow"))

# Age group
models_age = by(dataset_long, dataset_long$age_group, function(x) survival::coxph(Surv(time_in, time_out, endpoint) ~ fi_transformed + strata(as.factor(XAgeGrp)) + strata(is_female) + strata(region_code) + household_income + marital_status_group + occupation + highest_education + sbp_mean + random_glucose + alcohol_category + smoking_group + poor_diet + fat_body_mass_kg, data = x, ties = "breslow"))

# Area
models_region = by(dataset_long, dataset_long$region_is_urban, function(x) survival::coxph(Surv(time_in, time_out, endpoint) ~ fi_transformed + strata(as.factor(XAgeGrp)) + strata(is_female) + strata(region_code) + household_income + marital_status_group + occupation + highest_education + sbp_mean + random_glucose + alcohol_category + smoking_group + poor_diet + fat_body_mass_kg, data = x, ties = "breslow"))

# Education level
models_education = by(dataset_long, dataset_long$highest_education, function(x) survival::coxph(Surv(time_in, time_out, endpoint) ~ fi_transformed + strata(as.factor(XAgeGrp)) + strata(is_female) + strata(region_code) + household_income + marital_status_group + occupation + sbp_mean + random_glucose + alcohol_category + smoking_group + poor_diet + fat_body_mass_kg, data = x, ties = "breslow"))

# BMI 
models_fatmass = by(dataset_long, dataset_long$fat_body_mass_cat, function(x) survival::coxph(Surv(time_in, time_out, endpoint) ~ fi_transformed + strata(as.factor(XAgeGrp)) + strata(is_female) + strata(region_code) + household_income + marital_status_group + occupation + highest_education + sbp_mean + random_glucose + alcohol_category + smoking_group + poor_diet + fat_body_mass_kg, data = x, ties = "breslow"))

# Alcohol
models_alcohol = by(dataset_long, dataset_long$alcohol_category, function(x) survival::coxph(Surv(time_in, time_out, endpoint) ~ fi_transformed + strata(as.factor(XAgeGrp)) + strata(is_female) + strata(region_code) + household_income + marital_status_group + occupation + highest_education + sbp_mean + random_glucose + smoking_group + poor_diet + fat_body_mass_kg, data = x, ties = "breslow"))

# Smoking
models_smoking = by(dataset_long, dataset_long$smoking_group, function(x) survival::coxph(Surv(time_in, time_out, endpoint) ~ fi_transformed + strata(as.factor(XAgeGrp)) + strata(is_female) + strata(region_code) + household_income + marital_status_group + occupation + highest_education + sbp_mean + random_glucose + alcohol_category + poor_diet + fat_body_mass_kg, data = x, ties = "breslow"))

# Hypertension
models_sbp = by(dataset_long, dataset_long$sbp_group, function(x) survival::coxph(Surv(time_in, time_out, endpoint) ~ fi_transformed + strata(as.factor(XAgeGrp)) + strata(is_female) + strata(region_code) + household_income + marital_status_group + occupation + highest_education + sbp_mean + random_glucose + alcohol_category + smoking_group + poor_diet + fat_body_mass_kg, data = x, ties = "breslow"))

# Diabetes
models_glucose = by(dataset_long, dataset_long$random_glucose_group, function(x) survival::coxph(Surv(time_in, time_out, endpoint) ~ fi_transformed + strata(as.factor(XAgeGrp)) + strata(is_female) + strata(region_code) + household_income + marital_status_group + occupation + highest_education + sbp_mean + random_glucose + alcohol_category + smoking_group + poor_diet + fat_body_mass_kg, data = x, ties = "breslow"))
```

Write function to extract data from each model

```{r}
# Create function to extract data
extract_data = function(df, name, varlabel) {
  
  results = summary(df)$coefficients[1, c("coef", "se(coef)")]
  
  output = data.frame(
    #variable = paste0(varlabel, "_", names(x)),
    variable = paste0(varlabel, "_", name),
    estimate = results[["coef"]],
    stderr = results[["se(coef)"]],
    n = df$nevent
  )
  
  return(output)
  
}
```

Extract data for each subgroup analysis

```{r}
# Extract data for each of the subgroup and save as a list
data_sex = mapply(extract_data, df = models_sex, name = names(models_sex), varlabel="sex", SIMPLIFY = FALSE)
data_age = mapply(extract_data, df = models_age, name = names(models_age), varlabel="agegroup", SIMPLIFY = FALSE)
data_region = mapply(extract_data, df = models_region, name = names(models_region), varlabel="region", SIMPLIFY = FALSE)
data_education = mapply(extract_data, df = models_education, name = names(models_education), varlabel="education", SIMPLIFY = FALSE)
data_fatmass = mapply(extract_data, df = models_fatmass, name = names(models_fatmass), varlabel="fatmass", SIMPLIFY = FALSE)
data_alcohol = mapply(extract_data, df = models_alcohol, name = names(models_alcohol), varlabel="alcohol", SIMPLIFY = FALSE)
data_smoking = mapply(extract_data, df = models_smoking, name = names(models_smoking), varlabel="smoking", SIMPLIFY = FALSE)
data_sbp = mapply(extract_data, df = models_sbp, name = names(models_sbp), varlabel="sbp", SIMPLIFY = FALSE)
data_glucose = mapply(extract_data, df = models_glucose, name = names(models_glucose), varlabel="glucose", SIMPLIFY = FALSE)
```

Carry out heterogeneity and trend tests

```{r}
# Carry out heterogeneity or trend test
source("J:/R projects/PhD-by-chapter/source codes/ckb resources/trend_het.R")

# Heterogeneity
test_het_sex = heterogeneity(bind_rows(data_sex)$estimate, bind_rows(data_sex)$stderr)
test_het_region = heterogeneity(bind_rows(data_region)$estimate, bind_rows(data_region)$stderr)
test_het_alcohol = heterogeneity(bind_rows(data_alcohol)$estimate, bind_rows(data_alcohol)$stderr)
test_het_smoking = heterogeneity(bind_rows(data_smoking)$estimate, bind_rows(data_smoking)$stderr)
test_het_fatmass = heterogeneity(bind_rows(data_fatmass)$estimate, bind_rows(data_fatmass)$stderr)

# Trend
test_trend_age = trend(bind_rows(data_age)$estimate, bind_rows(data_age)$stderr)
test_trend_fatmass = trend(bind_rows(data_fatmass)$estimate, bind_rows(data_fatmass)$stderr)
test_trend_education = trend(bind_rows(data_education)$estimate, bind_rows(data_education)$stderr)
test_trend_sbp = heterogeneity(bind_rows(data_sbp)$estimate, bind_rows(data_sbp)$stderr)
test_trend_glucose = heterogeneity(bind_rows(data_glucose)$estimate, bind_rows(data_glucose)$stderr)
```

Combine output data into a clean dataset

```{r}
# Heterogeneity and trend tests
test_het = bind_rows(list("sex" = test_het_sex,
               "region" = test_het_region,
               "alcohol" = test_het_alcohol,
               "smoking" = test_het_smoking), .id = "variable")
names(test_het) = c("variable", "het_stat", "het_p")

test_trend = bind_rows(list("agegroup" = test_trend_age,
               "fatmass" = test_trend_fatmass,
               "education" = test_trend_education,
               "sbp" = test_trend_sbp,
               "glucose" = test_trend_glucose), .id = "variable")
names(test_trend) = c("variable", "trend_stat", "trend_p")

results_extra = full_join(test_het, test_trend, by = "variable") %>% 
  mutate(het_p = format.pval(het_p, eps=.001, digits=1),
         trend_p = format.pval(trend_p, eps=.001, digits=1))

# HRs
data_list = list(bind_rows(data_age),
                 bind_rows(data_sex),
                 bind_rows(data_region),
                 bind_rows(data_education),
                 bind_rows(data_fatmass),
                 bind_rows(data_alcohol),
                 bind_rows(data_smoking),
                 bind_rows(data_sbp),
                 bind_rows(data_glucose))

final_results = bind_rows(data_list)

final_results
```

```{r}
# Create group labels
final_results = final_results %>% 
  separate(variable, c("subheading", "label"), sep = "_", remove = FALSE) 

# Select the last group level of each subgroup to merge onto results_extra
labels_to_merge = final_results %>%
  group_by(subheading) %>% 
  dplyr::mutate(id = row_number()) %>% 
  filter(id == max(id)) %>% 
  select(variable, subheading)

results_extra = results_extra %>% 
  select(subheading = variable, everything()) %>% 
  full_join(labels_to_merge, by = "subheading") %>% # add labels
  select(-subheading) %>% 
  mutate_if(is.numeric, ~as.character(round(., 1))) %>% # round statistic to 1 dp
  mutate_at(vars(het_p, trend_p), ~if_else(grepl("^0", .), paste0("=", .), .)) # add = to pvalue

# Relabel headings 
final_results = final_results %>% 
  mutate(subheading = case_when(subheading == "agegroup" ~ "Age group",
                                subheading == "sex" ~ "Sex",
                                subheading == "region" ~ "Region",
                                subheading == "education" ~ "Highest education",
                                subheading == "fatmass" ~ "Fat mass quintile",
                                subheading == "alcohol" ~ "Alcohol category",
                                subheading == "smoking" ~ "Smoking category",
                                subheading == "sbp" ~ "Systolic blood pressure",
                                subheading == "glucose" ~ "Random glucose"))
```

Save data for plotting later

```{r}
stroke_results = list(final_results, results_extra)
ihd_results = list(final_results, results_extra)

#saveRDS(stroke_results, file = "J:/R projects/PhD-by-chapter/Chapter6-prospective-association/rds objects/output data/HRs baseline FI (modified) and stroke by subgroups.rds")
```

```{r, fig.height=10, fig.width=8}
row_labels = final_results[, c("subheading", "label", "variable")]
row_labels$heading = "Subgroup"

forestplot1 = ckbplotr::make_forest_plot(panels        = list(stroke_results[[1]]),
                           col.key       = "variable",
                           row.labels    = row_labels,
                           row.labels.levels = c("heading", "subheading", "label"),
                           rows          = "Subgroup",
                           exponentiate  = TRUE,
                           panel.names = c("Stroke"),
                           ci.delim      = "-",
                           xlim          = c(0.965, 1.21),
                           xticks        = c(1, 1.1, 1.2),
                           blankrows     = c(1, 1, 0, 1),
                           scalepoints   = TRUE,
                           pointsize     = 3,
                           col.left      = c("n"),
                           col.left.heading = c("No. of\nevents"),
                           col.heading.space = 1.5,
                           addtext       = list(stroke_results[[2]]),
                           right.space   = unit(35, "mm"))

forestplot2 = ckbplotr::make_forest_plot(panels        = list(ihd_results[[1]]),
                           col.key       = "variable",
                           row.labels    = row_labels,
                           row.labels.levels = c("heading", "subheading", "label"),
                           rows          = "Subgroup",
                           exponentiate  = TRUE,
                           panel.names = c("IHD"),
                           ci.delim      = "-",
                           xlim          = c(1, 1.265),
                           xticks        = c(1, 1.1, 1.2),
                           blankrows     = c(1, 1, 0, 1),
                           scalepoints   = TRUE,
                           pointsize     = 3,
                           col.left      = c("n"),
                           col.left.heading = c("No. of\nevents"),
                           col.heading.space = 1.5,
                           addtext       = list(ihd_results[[2]]),
                           right.space   = unit(35, "mm")) 
```

Save plots

```{r}
cowplot::save_plot("J:/R projects/PhD-by-chapter/Chapter6-prospective-association/outputs/age-at-risk stratified HRs/modified FI/HRplots_subgroups_FImodified_1SD_Stroke_fulladj.png", forestplot1$plot, base_height = 10, base_aspect_ratio = 0.7) 
cowplot::save_plot("J:/R projects/PhD-by-chapter/Chapter6-prospective-association/outputs/age-at-risk stratified HRs/modified FI/HRplots_subgroups_FImodified_1SD_IHD_fulladj.png", forestplot2$plot, base_height = 10, base_aspect_ratio = 0.7) 
```

Step 2: Remove the axis text for all but the first plot.

```{r}
p1 <- forestplot1$plot

p2 <- forestplot2$plot + 
  theme(axis.text.y = element_blank())
```

Step 3: Arrange the plots using gridExtra (there may be other packages that also work). Adjust widths until the plots are the desired widths in your final output file.

```{r, fig.height=10, fig.width=8}
final_plot = gridExtra::grid.arrange(p1, p2, nrow = 1, widths = c(1, 0.6))

cowplot::save_plot("J:/R projects/PhD-by-chapter/Chapter6-prospective-association/outputs/age-at-risk stratified HRs/modified FI/HRplots_subgroups_FImodified_1SD_IHDStroke_fulladj.png", final_plot, base_height = 10, base_aspect_ratio = 0.8) # aspect ratio = width/height
```

### Subgroup analysis for fat mass quintile

```{r}
# Heterogeneity and trend tests
test_het = as.data.frame(test_het_fatmass)
test_het$variable = "fatmass"
names(test_het) = c("het_stat", "het_p", "variable")

test_trend = as.data.frame(test_trend_fatmass)
test_trend$variable = "fatmass"
names(test_trend) = c("trend_stat", "trend_p", "variable")

results_extra = full_join(test_het, test_trend, by = "variable") %>% 
  mutate(het_p = format.pval(het_p, eps=.001, digits=1),
         trend_p = format.pval(trend_p, eps=.001, digits=1))

# HRs
final_results = bind_rows(data_fatmass)

final_results
```

```{r}
# Create group labels
final_results = final_results %>% 
  separate(variable, c("subheading", "label"), sep = "_", remove = FALSE) 

# Select the last group level of each subgroup to merge onto results_extra
labels_to_merge = final_results %>%
  group_by(subheading) %>% 
  dplyr::mutate(id = row_number()) %>% 
  filter(id == max(id)) %>% 
  select(variable, subheading)

results_extra = results_extra %>% 
  select(subheading = variable, everything()) %>% 
  full_join(labels_to_merge, by = "subheading") %>% # add labels
  select(-subheading) %>% 
  mutate_if(is.numeric, ~as.character(round(., 1))) %>% # round statistic to 1 dp
  mutate_at(vars(het_p, trend_p), ~if_else(grepl("^0", .), paste0("=", .), .)) # add = to pvalue

# Relabel headings 
final_results = final_results %>% 
  mutate(subheading = "Fat mass")
```

Save data for plotting later

```{r}
stroke_results = list(final_results, results_extra)
ihd_results = list(final_results, results_extra)

#saveRDS(stroke_results, file = "J:/R projects/PhD-by-chapter/Chapter6-prospective-association/rds objects/output data/HRs baseline FI (modified) and stroke by subgroups.rds")
```

```{r, fig.height=10, fig.width=8}
row_labels = final_results[, c("subheading", "label", "variable")]
row_labels$heading = "Subgroup"

forestplot1a = ckbplotr::make_forest_plot(panels        = list(stroke_results[[1]]),
                           col.key       = "variable",
                           row.labels    = row_labels,
                           row.labels.levels = c("heading", "subheading", "label"),
                           rows          = "Subgroup",
                           exponentiate  = TRUE,
                           panel.names = c("Stroke"),
                           ci.delim      = "-",
                           xlim          = c(1, 1.22),
                           xticks        = c(1, 1.1, 1.2),
                           blankrows     = c(1, 1, 0, 1),
                           scalepoints   = TRUE,
                           pointsize     = 3,
                           col.left      = c("n"),
                           col.left.heading = c("No. of\nevents"),
                           col.right.heading = c("HR (95% CI)\nper 1SD FI"),
                           col.heading.space = 1.5,
                           addtext       = list(stroke_results[[2]]),
                           right.space   = unit(35, "mm"))

forestplot2a = ckbplotr::make_forest_plot(panels        = list(ihd_results[[1]]),
                           col.key       = "variable",
                           row.labels    = row_labels,
                           row.labels.levels = c("heading", "subheading", "label"),
                           rows          = "Subgroup",
                           exponentiate  = TRUE,
                           panel.names = c("IHD"),
                           ci.delim      = "-",
                           xlim          = c(1, 1.22),
                           xticks        = c(1, 1.1, 1.2),
                           blankrows     = c(1, 1, 0, 1),
                           scalepoints   = TRUE,
                           pointsize     = 3,
                           col.left      = c("n"),
                           col.left.heading = c("No. of\nevents"),
                           col.right.heading = c("HR (95% CI)\nper 1SD FI"),
                           col.heading.space = 1.5,
                           addtext       = list(ihd_results[[2]]),
                           right.space   = unit(35, "mm")) 
```

Merge and save plot

```{r, fig.height=5, fig.width=7}
# Step 2: Remove the axis text for all but the first plot
p1 <- forestplot1a$plot

p2 <- forestplot2a$plot + 
  theme(axis.text.y = element_blank())

# Step 3: Arrange the plots using gridExtra (there may be other packages that also work). Adjust widths until the plots are the desired widths in your final output file
final_plot = gridExtra::grid.arrange(p1, p2, nrow = 1, widths = c(1, 0.7))

cowplot::save_plot("J:/R projects/PhD-by-chapter/Chapter6-prospective-association/outputs/age-at-risk stratified HRs/modified FI/HRplots_subgroups_FImodified_1SD_IHDStroke_fulladj_fatmass.png", final_plot, base_height = 5, base_aspect_ratio = 2) # aspect ratio = width/height
```
