---
title: "CoxPH subgroup models"
output: html_document
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Required libraries

```{r library}
library(broom)      # for tidy() to get clean HRs and confidence intervals from model fit
library(cowplot)
library(data.table)
library(Epi)        # for float(), which compute floating absolute risks using the method by Easton et al.
library(gridExtra)
library(kableExtra)
library(knitr)
library(lubridate)
library(survival)   # for Surv(), cox regression
library(tidyverse)
```

## Open and process data

Run external script to open and process data

```{r open-data}
# Source script that creates working dataset that includes baseline frailty scores and CVD outcomes with key baseline covariates
source("J:/R projects/PhD-by-chapter/Chapter6-prospective-association/code/source/script to open and process dataset to analyse baseline frailty and outcomes.R")

# open pretty covariates for later
covariates = readRDS("J:/R projects/PhD-by-chapter/rds objects/covariates_pretty_levels_for_everyone.rds")
```

Data structure

```{r data-structure}
str(mydata)
```


Number of participants included:

  - FI analysis: `r nrow(mydata)`

Further process data for age-at-risk analysis to have following variables:

  - csid
  - dob_anon
  - study_date
  - endpoint - binary variable indicating the endpoint (1/0)
  - endpoint_date - date of endpoint event or censoring
  - frailty quintile group
  - outcome_type - indicate outcome type, e.g. stroke, MI, etc.

```{r data-process, message=FALSE, warning=FALSE}
mydata = mydata %>% 
  mutate_at(vars(contains("_ind")), ~replace(., is.na(.), 0)) # replace missing values with zeros for dummy variables 

mydata = mydata %>% 
  mutate_at(vars(contains("total_date"), contains("incidence_date"), contains("fatal_date"), date_of_death), ~replace(., is.na(.), censoring_date)) # if empty dates (i.e. no event), then replace empty values with censoring date

kable(head(mydata)) %>% kableExtra::kable_styling()
```

Transform FI variable

```{r fi-process}
if (fi_type == "mFI-1"){
  
  mydata$fi_transformed = (mydata$fi.score_no_cv - mean(mydata$fi.score_no_cv))/sd(mydata$fi.score_no_cv) # per 1 SD
  
} else if (fi_type == "mFI-3"){
  
  mydata$fi_transformed = (mydata$fi.score_no_21 - mean(mydata$fi.score_no_21))/sd(mydata$fi.score_no_21) # per 1 SD
  
} else {}

#mydata$fi_deficit = mydata$fi.score_no_cv/(1/23) # per 1 deficit increment
#mydata$fi_0.1 = mydata$fi.score_no_cv/0.1 # per 0.1 unit increment

#hist(mydata$fi_transformed)
#hist(mydata$fi.score_no_cv)
```

Make outcome data long

```{r make-data-long}
temp = mydata %>% 
  select(csid, event_All_cause_mortality_ind = died, event_All_cause_mortality_date = date_of_death, contains("event_")) # rename death variables

temp = pivot_longer(temp, cols = -1, names_pattern = "(.*)(ind|date)$", names_to = c("names", ".value"))
  # names_pattern = ()()  means look for 2 parts; (.*) means first part should have zero or more characters; (ind|date)$ means the second part should end either as ind or date

temp$names = as.factor(temp$names)
```

Split data on outcome and save as list

```{r split-data}
mylist = split(temp, f = temp$names)
```

## Analysis

Select dataset with outcome

```{r}
dataset = mylist[[c(outcome)]]
```

Expand data for age-at-risk analysis

```{r}
# Expand data for age-at-risk analysis
source("J:/R projects/PhD-by-chapter/source codes/ckb resources/expand_age_at_risk.R")

# Merge dob_anon, study_date and rename variables
dataset = dataset %>% 
  left_join(select(mydata, csid, dob_anon, study_date), by = "csid") %>% # mydata contains frailty variable and other covariates
  select(csid, dob_anon, study_date, endpoint = ind, endpoint_date = date)

dataset_long = expand_age_at_risk(df = dataset,
                                  ages = c(40, 45, 50, 55, 60, 65, 70, 75, 80))
#head(dataset_long)

# Merge columns - expanded dataset does not contain the risk factor (frailty) or other covariates
to_merge = covariates %>% 
  select(csid, is_female, age_at_study_date, age_group, region_is_urban, region_code, household_income, marital_status_group, occupation, highest_education, met, hypertension_diag, diabetes_diag, random_glucose_group, sbp_group, fat_body_mass_kg, alcohol_category, smoking_group, smoking_group_qty, poor_diet, sbp_mean, random_glucose) %>% 
  mutate(age_group2 = factor(case_when(age_at_study_date < 50 ~ "30-49",
                                       age_at_study_date < 55 ~ "50-54",
                                       age_at_study_date < 60 ~ "55-59",
                                       age_at_study_date < 65 ~ "60-64",
                                       age_at_study_date < 70 ~ "65-69",
                                       TRUE ~ "70-80"), levels = c("30-49", "50-54", "55-59", "60-64", "65-69", "70-80")),
         highest_education_cat = factor(case_when(highest_education == "No formal school" ~ "No formal school",
                                                  highest_education == "Primary School" ~ "Primary school",
                                                  TRUE ~ "Middle school or higher"), levels = c("No formal school", "Primary school", "Middle school or higher")),
         fat_body_mass_cat = factor(ntile(fat_body_mass_kg, 5), levels=c(1:5)),
         sbp_cat = factor(ntile(sbp_mean, 5), levels=c(1:5)),
         random_glucose_cat = factor(ntile(random_glucose, 5), levels=c(1:5))
         )# recode variables
  
to_merge = to_merge %>% 
  group_by(is_female) %>% # sex-specific MET groups
  mutate(met_cat = factor(ntile(met, 5), levels=c(1:5)) ) %>% 
  ungroup()

dataset_long = left_join(dataset_long, to_merge, by = "csid") %>% 
  left_join(select(mydata, csid, fi_transformed), by = "csid")
```


Run age-at-risk, sex, and region stratified, where appropriate, Cox proportional hazards models by subgroup

```{r}
# Sex
models_sex = by(dataset_long, dataset_long$is_female, function(x) survival::coxph(Surv(time_in, time_out, endpoint) ~ fi_transformed + strata(as.factor(XAgeGrp)) + strata(region_code) + household_income + marital_status_group + occupation + highest_education + sbp_mean + random_glucose + alcohol_category + smoking_group + poor_diet + fat_body_mass_kg + met, data = x, ties = "breslow"))

# Age group
models_age = by(dataset_long, dataset_long$age_group, function(x) survival::coxph(Surv(time_in, time_out, endpoint) ~ fi_transformed + strata(as.factor(XAgeGrp)) + strata(is_female) + strata(region_code) + household_income + marital_status_group + occupation + highest_education + sbp_mean + random_glucose + alcohol_category + smoking_group + poor_diet + fat_body_mass_kg + met, data = x, ties = "breslow"))

models_age2 = by(dataset_long, dataset_long$age_group2, function(x) survival::coxph(Surv(time_in, time_out, endpoint) ~ fi_transformed + strata(as.factor(XAgeGrp)) + strata(is_female) + strata(region_code) + household_income + marital_status_group + occupation + highest_education + sbp_mean + random_glucose + alcohol_category + smoking_group + poor_diet + fat_body_mass_kg + met, data = x, ties = "breslow"))

# Area
models_region = by(dataset_long, dataset_long$region_is_urban, function(x) survival::coxph(Surv(time_in, time_out, endpoint) ~ fi_transformed + strata(as.factor(XAgeGrp)) + strata(is_female) + strata(region_code) + household_income + marital_status_group + occupation + highest_education + sbp_mean + random_glucose + alcohol_category + smoking_group + poor_diet + fat_body_mass_kg + met, data = x, ties = "breslow"))

models_region2 = by(dataset_long, dataset_long$region_code, function(x) survival::coxph(Surv(time_in, time_out, endpoint) ~ fi_transformed + strata(as.factor(XAgeGrp)) + strata(is_female) + household_income + marital_status_group + occupation + highest_education + sbp_mean + random_glucose + alcohol_category + smoking_group + poor_diet + fat_body_mass_kg + met, data = x, ties = "breslow"))

# Education level
models_education = by(dataset_long, dataset_long$highest_education, function(x) survival::coxph(Surv(time_in, time_out, endpoint) ~ fi_transformed + strata(as.factor(XAgeGrp)) + strata(is_female) + strata(region_code) + household_income + marital_status_group + occupation + sbp_mean + random_glucose + alcohol_category + smoking_group + poor_diet + fat_body_mass_kg + met, data = x, ties = "breslow"))

models_education2 = by(dataset_long, dataset_long$highest_education_cat, function(x) survival::coxph(Surv(time_in, time_out, endpoint) ~ fi_transformed + strata(as.factor(XAgeGrp)) + strata(is_female) + strata(region_code) + household_income + marital_status_group + occupation + sbp_mean + random_glucose + alcohol_category + smoking_group + poor_diet + fat_body_mass_kg + met, data = x, ties = "breslow"))

# Household income
models_income = by(dataset_long, dataset_long$household_income, function(x) survival::coxph(Surv(time_in, time_out, endpoint) ~ fi_transformed + strata(as.factor(XAgeGrp)) + strata(is_female) + strata(region_code) + marital_status_group + occupation + sbp_mean + random_glucose + alcohol_category + smoking_group + poor_diet + fat_body_mass_kg + met, data = x, ties = "breslow"))

# Fat mass
models_fatmass = by(dataset_long, dataset_long$fat_body_mass_cat, function(x) survival::coxph(Surv(time_in, time_out, endpoint) ~ fi_transformed + strata(as.factor(XAgeGrp)) + strata(is_female) + strata(region_code) + household_income + marital_status_group + occupation + highest_education + sbp_mean + random_glucose + alcohol_category + smoking_group + poor_diet + met, data = x, ties = "breslow"))

# Alcohol
models_alcohol = by(dataset_long, dataset_long$alcohol_category, function(x) survival::coxph(Surv(time_in, time_out, endpoint) ~ fi_transformed + strata(as.factor(XAgeGrp)) + strata(is_female) + strata(region_code) + household_income + marital_status_group + occupation + highest_education + sbp_mean + random_glucose + smoking_group + poor_diet + fat_body_mass_kg + met, data = x, ties = "breslow"))

# Smoking
models_smoking = by(dataset_long, dataset_long$smoking_group, function(x) survival::coxph(Surv(time_in, time_out, endpoint) ~ fi_transformed + strata(as.factor(XAgeGrp)) + strata(is_female) + strata(region_code) + household_income + marital_status_group + occupation + highest_education + sbp_mean + random_glucose + alcohol_category + poor_diet + fat_body_mass_kg + met, data = x, ties = "breslow"))

models_smoking2 = by(dataset_long, dataset_long$smoking_group_qty, function(x) survival::coxph(Surv(time_in, time_out, endpoint) ~ fi_transformed + strata(as.factor(XAgeGrp)) + strata(is_female) + strata(region_code) + household_income + marital_status_group + occupation + highest_education + sbp_mean + random_glucose + alcohol_category + poor_diet + fat_body_mass_kg + met, data = x, ties = "breslow"))

# Hypertension
models_hypertension = by(dataset_long, dataset_long$hypertension_diag, function(x) survival::coxph(Surv(time_in, time_out, endpoint) ~ fi_transformed + strata(as.factor(XAgeGrp)) + strata(is_female) + strata(region_code) + household_income + marital_status_group + occupation + highest_education + sbp_mean + random_glucose + alcohol_category + smoking_group + poor_diet + fat_body_mass_kg + met, data = x, ties = "breslow"))

models_sbp = by(dataset_long, dataset_long$sbp_group, function(x) survival::coxph(Surv(time_in, time_out, endpoint) ~ fi_transformed + strata(as.factor(XAgeGrp)) + strata(is_female) + strata(region_code) + household_income + marital_status_group + occupation + highest_education + random_glucose + alcohol_category + smoking_group + poor_diet + fat_body_mass_kg + met, data = x, ties = "breslow"))

models_sbp2 = by(dataset_long, dataset_long$sbp_cat, function(x) survival::coxph(Surv(time_in, time_out, endpoint) ~ fi_transformed + strata(as.factor(XAgeGrp)) + strata(is_female) + strata(region_code) + household_income + marital_status_group + occupation + highest_education + random_glucose + alcohol_category + smoking_group + poor_diet + fat_body_mass_kg + met, data = x, ties = "breslow"))

# Diabetes
models_diabetes = by(dataset_long, dataset_long$diabetes_diag, function(x) survival::coxph(Surv(time_in, time_out, endpoint) ~ fi_transformed + strata(as.factor(XAgeGrp)) + strata(is_female) + strata(region_code) + household_income + marital_status_group + occupation + highest_education + sbp_mean + random_glucose + alcohol_category + smoking_group + poor_diet + fat_body_mass_kg + met, data = x, ties = "breslow"))

models_glucose = by(dataset_long, dataset_long$random_glucose_group, function(x) survival::coxph(Surv(time_in, time_out, endpoint) ~ fi_transformed + strata(as.factor(XAgeGrp)) + strata(is_female) + strata(region_code) + household_income + marital_status_group + occupation + highest_education + sbp_mean  + alcohol_category + smoking_group + poor_diet + fat_body_mass_kg + met, data = x, ties = "breslow"))

models_glucose2 = by(dataset_long, dataset_long$random_glucose_cat, function(x) survival::coxph(Surv(time_in, time_out, endpoint) ~ fi_transformed + strata(as.factor(XAgeGrp)) + strata(is_female) + strata(region_code) + household_income + marital_status_group + occupation + highest_education + sbp_mean  + alcohol_category + smoking_group + poor_diet + fat_body_mass_kg + met, data = x, ties = "breslow"))

# Physical activity
models_met = by(dataset_long, dataset_long$met_cat, function(x) survival::coxph(Surv(time_in, time_out, endpoint) ~ fi_transformed + strata(as.factor(XAgeGrp)) + strata(is_female) + strata(region_code) + household_income + marital_status_group + occupation + highest_education + sbp_mean + random_glucose + alcohol_category + smoking_group + poor_diet + fat_body_mass_kg, data = x, ties = "breslow"))
```

Write function to extract data from each model

```{r}
# Create function to extract data
extract_data = function(df, name, varlabel) {
  
  results = summary(df)$coefficients[1, c("coef", "se(coef)")]
  
  output = data.frame(
    #variable = paste0(varlabel, "_", names(x)),
    variable = paste0(varlabel, "_", name),
    estimate = results[["coef"]],
    stderr = results[["se(coef)"]],
    n = df$nevent
  )
  
  return(output)
  
}
```

Extract data for each subgroup analysis

```{r}
# Extract data for each of the subgroup and save as a list
sub_data_sex = mapply(extract_data, df = models_sex, name = names(models_sex), varlabel="sex", SIMPLIFY = FALSE)
sub_data_age = mapply(extract_data, df = models_age, name = names(models_age), varlabel="agegroup", SIMPLIFY = FALSE)
sub_data_age2 = mapply(extract_data, df = models_age2, name = names(models_age2), varlabel="agegroup2", SIMPLIFY = FALSE)
sub_data_region = mapply(extract_data, df = models_region, name = names(models_region), varlabel="region", SIMPLIFY = FALSE)
sub_data_region2 = mapply(extract_data, df = models_region2, name = names(models_region2), varlabel="region2", SIMPLIFY = FALSE)
sub_data_income = mapply(extract_data, df = models_income, name = names(models_income), varlabel="income", SIMPLIFY = FALSE)
sub_data_education = mapply(extract_data, df = models_education, name = names(models_education), varlabel="education", SIMPLIFY = FALSE)
sub_data_education2 = mapply(extract_data, df = models_education2, name = names(models_education2), varlabel="education2", SIMPLIFY = FALSE)
sub_data_fatmass = mapply(extract_data, df = models_fatmass, name = names(models_fatmass), varlabel="fatmass", SIMPLIFY = FALSE)
sub_data_alcohol = mapply(extract_data, df = models_alcohol, name = names(models_alcohol), varlabel="alcohol", SIMPLIFY = FALSE)
sub_data_smoking = mapply(extract_data, df = models_smoking, name = names(models_smoking), varlabel="smoking", SIMPLIFY = FALSE)
sub_data_smoking2 = mapply(extract_data, df = models_smoking2, name = names(models_smoking2), varlabel="smoking2", SIMPLIFY = FALSE)
sub_data_hypertension = mapply(extract_data, df = models_hypertension, name = names(models_hypertension), varlabel="hypertension", SIMPLIFY = FALSE)
sub_data_sbp = mapply(extract_data, df = models_sbp, name = names(models_sbp), varlabel="sbp", SIMPLIFY = FALSE)
sub_data_sbp2 = mapply(extract_data, df = models_sbp2, name = names(models_sbp2), varlabel="sbp2", SIMPLIFY = FALSE)
sub_data_diabetes = mapply(extract_data, df = models_diabetes, name = names(models_diabetes), varlabel="diabetes", SIMPLIFY = FALSE)
sub_data_glucose = mapply(extract_data, df = models_glucose, name = names(models_glucose), varlabel="glucose", SIMPLIFY = FALSE)
sub_data_glucose2 = mapply(extract_data, df = models_glucose2, name = names(models_glucose2), varlabel="glucose2", SIMPLIFY = FALSE)
sub_data_met = mapply(extract_data, df = models_met, name = names(models_met), varlabel="met", SIMPLIFY = FALSE)
```

Carry out heterogeneity and trend tests

```{r}
# Carry out heterogeneity or trend test
source("J:/R projects/PhD-by-chapter/source codes/ckb resources/trend_het.R")

# Heterogeneity test
run_het_test = function(df){
  
  test = heterogeneity(bind_rows(df)$estimate, bind_rows(df)$stderr)
  
  return(test)
}

# Trend test
run_trend_test = function(df){
  
  test = trend(bind_rows(df)$estimate, bind_rows(df)$stderr)
  
  return(test)
}

# Load all data frames in current environment that begins with 'sub_data_' into a list
dfs = lapply(ls(pattern="^sub_data_"), function(x) get(x))

# Run tests in each element of list
het_ls = lapply(dfs, run_het_test)
trend_ls = lapply(dfs, run_trend_test)

# Name the list elements
dfs_name = ls(pattern ="^sub_data_")
names(het_ls) = dfs_name
names(trend_ls) = dfs_name
```

Combine output data into a clean dataset

```{r}
# Heterogeneity and trend tests
test_het = bind_rows(het_ls, .id = "rf")
names(test_het) = c("rf", "het_stat", "het_p")

test_trend = bind_rows(trend_ls, .id = "rf")
names(test_trend) = c("rf", "trend_stat", "trend_p")

results_extra = full_join(test_het, test_trend, by = "rf") %>% 
  mutate(het_p = format.pval(het_p, eps=.001, digits=1),
         trend_p = format.pval(trend_p, eps=.001, digits=1))

# HRs
data_list = lapply(dfs, bind_rows) # for each element in list, bind rows
names(data_list) = dfs_name # name each list element

final_results = bind_rows(data_list, .id = "rf") # bind list elements into a single data frame
```

```{r}
# Create group labels
final_results = final_results %>% 
  separate(variable, c("subheading", "label"), sep = "_", remove = FALSE) 

# Select the last group level of each subgroup to merge onto results_extra
labels_to_merge = final_results %>%
  group_by(subheading) %>% 
  dplyr::mutate(id = row_number()) %>% 
  filter(id == max(id)) %>% 
  select(rf, label)

results_extra = results_extra %>%
  full_join(labels_to_merge, by = "rf") %>% # add labels
  mutate_if(is.numeric, ~as.character(round(., 1))) %>% # round statistic to 1 dp
  mutate_at(vars(het_p, trend_p), ~if_else(grepl("^0", .), paste0("=", .), .)) # add = to pvalue

# Relabel headings 
final_results = final_results %>% 
  mutate(subheading = case_when(subheading == "agegroup" ~ "Age group",
                                subheading == "agegroup2" ~ "Age groups",
                                subheading == "sex" ~ "Sex",
                                subheading == "region" ~ "Region",
                                subheading == "region2" ~ "Study site",
                                subheading == "education" ~ "Highest education level",
                                subheading == "education2" ~ "Highest education",
                                subheading == "income" ~ "Income level",
                                subheading == "fatmass" ~ "Fat mass quintile",
                                subheading == "alcohol" ~ "Alcohol category",
                                subheading == "smoking" ~ "Smoking category",
                                subheading == "smoking2" ~ "Smoking category (quantity)",
                                subheading == "hypertension" ~ "Hypertension",
                                subheading == "diabetes" ~ "Diabetes",
                                subheading == "sbp" ~ "Systolic blood pressure",
                                subheading == "glucose" ~ "Random glucose",
                                subheading == "sbp2" ~ "SBP quintile",
                                subheading == "glucose2" ~ "Random glucose quintile",
                                subheading == "met" ~ "Physical activity quintile"))

final_results
results_extra
```

Save data for plotting later

```{r}
save_results = list(final_results, results_extra)

saveRDS(save_results, file = paste0("J:/R projects/PhD-by-chapter/Chapter6-prospective-association/rds objects/output data/subgroup analysis/HRs ", fi_type, " and ", outcome, "subgroup.rds"))
```


```{r}
knitr::knit_exit()
```
