---
title: "Test cox model assumption"
output: html_document
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Cox Model Assumptions

 - Proportional hazards assumptions can be checked by *Schoenfeld residuals*
 - Nonlinearity in relationship between the log hazard and the covariates can be checked by *Martingale residual*
 - Influential observations or outliers can be checked by *deviance residual* (symmetric transformation of the Maringale residuals)
 
#### Required libraries

```{r libraries}
library(gridExtra)
library(survminer)

# required for age-at-risk stratified CoxPH model
library(Epi)
library(lubridate)
library(survival)   
library(tidyverse)
```

#### Open and process data

```{r open-data}
# Source script that creates working dataset that includes baseline frailty scores and CVD outcomes with key baseline covariates
source("J:/R projects/PhD-by-chapter/Chapter6-prospective-association/code/source/script to open and process dataset to analyse baseline frailty and outcomes.R")

mydata = mydata %>% 
  mutate(across(contains("_ind"), ~replace(., is.na(.), 0))) # replace missing values with zeros for dummy variables 

mydata = mydata %>% 
  mutate(across(c(contains("total_date"), contains("incidence_date"), contains("fatal_date"), date_of_death), ~replace(., is.na(.), censoring_date))) # if empty dates (i.e. no event), then replace empty values with censoring date

mydata$fi5= factor(ntile(mydata$fi.score_no_cv, 5)) # split FI into quintile groups 
mydata$ficat = mydata$fi.cat_no_cv
mydata$fi_transformed= (mydata$fi.score_no_cv - mean(mydata$fi.score_no_cv))/sd(mydata$fi.score_no_cv) # per 1 SD

temp = mydata %>% 
  select(csid, event_All_cause_mortality_ind = died, event_All_cause_mortality_date = date_of_death, contains("Stroke_total"), contains("IHD_total"), contains("Fatal_CVD_total"))

temp = pivot_longer(temp, cols = -1, names_pattern = "(.*)(ind|date)$", names_to = c("names", ".value"))
  # names_pattern = ()()  means look for 2 parts; (.*) means first part should have zero or more characters; (ind|date)$ means the second part should end either as ind or date

temp$names = as.factor(temp$names)

mylist = split(temp, f = temp$names)
```

```{r}
names(mylist)
```

### Test proportional hazards assumptions

Test PH assumption between *transformed mFI-1* and *total stroke, total IHD, and CVD mortality.*

A plot that shows a non-random pattern of the Schoenfeld residuals against time is evidence of violation of the PH assumption

Use `cox.zph()` function in the `survival` package to test PH assumption for each covariate included in a Cox regression model fit. It additionally performs a global test for the model as a whole.

#### Total stroke

Compute a Cox regression model

```{r}
source("J:/R projects/PhD-by-chapter/source codes/ckb resources/expand_age_at_risk.R")

  # Merge dob_anon, study_date and rename variables
dataset = mylist[[6]] %>% 
  left_join(select(mydata, csid, dob_anon, study_date, fi_transformed, is_female), by = "csid") %>% # mydata contains frailty variable and other covariates
  select(csid, dob_anon, study_date, endpoint = ind, endpoint_date = date, fi_transformed, is_female)
  
formula = as.formula(paste("Surv(time_in, time_out, endpoint) ~ fi_transformed + strata(as.factor(XAgeGrp)) + is_female + region_code + household_income + marital_status + occupation + highest_education + hypertension_diag + diabetes_diag + alcohol_category + smoking_group + fat_body_mass_kg + poor_diet + met"))
    
# Expand data by age-at-risk
dataset_long = expand_age_at_risk(df = dataset,
                                  ages = c(40, 45, 50, 55, 60, 65, 70, 75, 80))
  
# Merge columns - expanded dataset does not contain the risk factor (frailty) or other covariates
to_merge = mydata %>% 
  select(csid, fi_transformed, is_female, region_code, household_income, marital_status, occupation, highest_education, met, hypertension_diag, diabetes_diag, alcohol_category, smoking_group, fat_body_mass_kg, poor_diet)
  
dataset_long = left_join(dataset_long, to_merge, by = "csid")
  
# Cox proportional hazards model
fit_mfi1_stroke = survival::coxph(formula, data = dataset_long, ties = "breslow") # the expanded dataframe contains 'XAgeGrp' column which can be used to stratify the CoxPH model, the time columns specify the start and end time for intervals
  #summary(fit)
```

Test PH assumption

```{r}
test.ph = survival::cox.zph(fit_mfi1_stroke)
test.ph
```

This is a goodness-of-fit test, which tests the correlation between the Schoenfeld residuals and survival time. A correlation of zero indicates that the model met the PH assumption (the null hypothesis).

The PH assumption is not satisfied for all covariates except for FI and poor diet variable.

Graph of the scaled Schoenfeld residuals against the transformed time

```{r, out.width="100%"}
test.ph.mfi1.stroke = survminer::ggcoxzph(test.ph)

test.ph.mfi1.stroke

ggplot2::ggsave("J:/R projects/PhD-by-chapter/Chapter6-prospective-association/outputs/assumptions/testPH-mFi1-totalstroke.png", width=45, height=45, units="cm", gridExtra::arrangeGrob(grobs = test.ph.mfi1.stroke))
```

#### Total IHD

Compute a Cox regression model

```{r}
# Merge dob_anon, study_date and rename variables
dataset = mylist[[4]] %>% 
  left_join(select(mydata, csid, dob_anon, study_date, fi_transformed, is_female), by = "csid") %>% # mydata contains frailty variable and other covariates
  select(csid, dob_anon, study_date, endpoint = ind, endpoint_date = date, fi_transformed, is_female)
  
formula = as.formula(paste("Surv(time_in, time_out, endpoint) ~ fi_transformed + strata(as.factor(XAgeGrp)) + is_female + region_code + household_income + marital_status + occupation + highest_education + hypertension_diag + diabetes_diag + alcohol_category + smoking_group + fat_body_mass_kg + poor_diet + met"))
    
# Expand data by age-at-risk
dataset_long = expand_age_at_risk(df = dataset,
                                  ages = c(40, 45, 50, 55, 60, 65, 70, 75, 80))
  
# Merge columns - expanded dataset does not contain the risk factor (frailty) or other covariates
to_merge = mydata %>% 
  select(csid, fi_transformed, is_female, region_code, household_income, marital_status, occupation, highest_education, met, hypertension_diag, diabetes_diag, alcohol_category, smoking_group, fat_body_mass_kg, poor_diet)
  
dataset_long = left_join(dataset_long, to_merge, by = "csid")
  
# Cox proportional hazards model
fit_mfi1_ihd = survival::coxph(formula, data = dataset_long, ties = "breslow") # the expanded dataframe contains 'XAgeGrp' column which can be used to stratify the CoxPH model, the time columns specify the start and end time for intervals
  #summary(fit)
```

Test PH assumption

```{r}
test.ph.ihd = survival::cox.zph(fit_mfi1_ihd)
test.ph.ihd
```

This is a goodness-of-fit test, which tests the correlation between the Schoenfeld residuals and survival time. A correlation of zero indicates that the model met the PH assumption (the null hypothesis).

The PH assumption is not satisfied for all covariates except for FI and poor diet variable.

Graph of the scaled Schoenfeld residuals against the transformed time

```{r, out.width="100%"}
test.ph.mfi1.ihd = survminer::ggcoxzph(test.ph.ihd)
test.ph.mfi1.ihd 

ggplot2::ggsave("J:/R projects/PhD-by-chapter/Chapter6-prospective-association/outputs/assumptions/testPH-mFi1-totalihd.png", width=45, height=45, units="cm", gridExtra::arrangeGrob(grobs = test.ph.mfi1.ihd))
```

#### Total CVD mortality

Compute a Cox regression model

```{r}
# Merge dob_anon, study_date and rename variables
dataset = mylist[[2]] %>% 
  left_join(select(mydata, csid, dob_anon, study_date, fi_transformed, is_female), by = "csid") %>% # mydata contains frailty variable and other covariates
  select(csid, dob_anon, study_date, endpoint = ind, endpoint_date = date, fi_transformed, is_female)
  
formula = as.formula(paste("Surv(time_in, time_out, endpoint) ~ fi_transformed + strata(as.factor(XAgeGrp)) + is_female + region_code + household_income + marital_status + occupation + highest_education + hypertension_diag + diabetes_diag + alcohol_category + smoking_group + fat_body_mass_kg + poor_diet + met"))
    
# Expand data by age-at-risk
dataset_long = expand_age_at_risk(df = dataset,
                                  ages = c(40, 45, 50, 55, 60, 65, 70, 75, 80))
  
# Merge columns - expanded dataset does not contain the risk factor (frailty) or other covariates
to_merge = mydata %>% 
  select(csid, fi_transformed, is_female, region_code, household_income, marital_status, occupation, highest_education, met, hypertension_diag, diabetes_diag, alcohol_category, smoking_group, fat_body_mass_kg, poor_diet)
  
dataset_long = left_join(dataset_long, to_merge, by = "csid")
  
# Cox proportional hazards model
fit_mfi1_cvd = survival::coxph(formula, data = dataset_long, ties = "breslow") # the expanded dataframe contains 'XAgeGrp' column which can be used to stratify the CoxPH model, the time columns specify the start and end time for intervals
  #summary(fit)
```

Test PH assumption

```{r}
test.ph.cvd = survival::cox.zph(fit_mfi1_cvd)
test.ph.cvd
```

This is a goodness-of-fit test, which tests the correlation between the Schoenfeld residuals and survival time. A correlation of zero indicates that the model met the PH assumption (the null hypothesis).

The PH assumption is not satisfied for all covariates except for FI and poor diet variable.

Graph of the scaled Schoenfeld residuals against the transformed time

```{r, out.width="100%"}
test.ph.mfi1.cvd = survminer::ggcoxzph(test.ph.cvd)
test.ph.mfi1.cvd 

ggplot2::ggsave("J:/R projects/PhD-by-chapter/Chapter6-prospective-association/outputs/assumptions/testPH-mFi1-totalcvd.png", width=45, height=45, units="cm", gridExtra::arrangeGrob(grobs = test.ph.mfi1.cvd))
```